### 2022-07-24 analysis/index.qmd
Use a Quarto website

---
title: "Analysis notebook"
---

This is a Quarto website. Neato.