### 2022-07-24 analysis/targets.qmd
Show overview of targets workflow

---
title: "Targets workflow"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

# targets pipeline

We use [the magical **`targets`** package](https://docs.ropensci.org/targets/) to run our analysis and keep track of all dependencies automatically.

To build our entire project, run `targets::tar_make()` at the R console.

Here's our complete pipeline:

```{r show-targets-pipeline, echo=FALSE}
withr::with_dir(here::here(), {
  targets::tar_glimpse()
})
```

# Actual code

All the data processing is handled with dataset-specific functions that live in `R/funs_data-cleaning.R`, which **`targets`** then runs as needed. For the sake of transparency, here's that code:

```{r generate-code-chunks, echo=FALSE}
# MAGIC: https://gist.github.com/StevenMMortimer/e54ec050d97d79996189
generate_chunk <- function(filename) {
  paste0(c(paste0("#### `R/", filename, "`"),
           paste0('```{r, code=xfun::read_utf8(here::here("R", "', filename, '")), eval=FALSE}'), 
           "#| code-fold: true", 
           "```", "", ""), 
         sep = "\n")
}

out <- NULL
for (thing in list.files(here::here("R"))) {
  out <- c(out, generate_chunk(thing))
}
```

`r paste(knitr::knit(text = out), collapse = "\n")`