### 2022-07-24 analysis/data-details.qmd
Add Oxford section

---
title: "Data details"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "png", dev.args = list(type = "cairo-png"))

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(DT)
library(kableExtra)
library(lubridate)

matrix_from_vector <- function(x, ncol) {
  n_balanced <- ceiling(length(x) / ncol) * ncol
  matrix(c(x, rep(NA, n_balanced - length(x))), ncol = ncol)
}

# Generated via random.org
set.seed(2385)

# Load data 
withr::with_dir(here::here(), {
  # Final data
  tar_load(daily_panel)
})
```

```{r extract-details}
info <- list(
  n_days = length(unique(daily_panel$day)),
  n_countries = length(unique(daily_panel$iso3)),
  n_rows = scales::label_comma()(nrow(daily_panel)),
  dates = map(list(
    start_date = min(daily_panel$day),
    end_date = max(daily_panel$day)
  ), ~list(date = ., 
           nice = format(., "%B %e, %Y"),
           year = year(.)))
)
```

## Overview

We combine data from the [International Covenant on Civil and Political Rights](https://www.ohchr.org/en/instruments-mechanisms/instruments/international-covenant-civil-and-political-rights) (ICCPR), the [World Health Organization](https://covid19.who.int/data) (WHO), [Oxford's COVID-19 Government Response Tracker](https://www.bsg.ox.ac.uk/research/research-projects/covid-19-government-response-tracker), and the [Varieties of Democracy](https://www.v-dem.net/) (V-Dem) project to track government responses, COVID case counts, and derogation status at a daily level.

### Countries and dates included

We include data from `r info$n_countries` countries for `r info$n_days` days from `r info$dates$start_date$nice` to `r info$dates$end_date$nice`, resulting in a panel with `r info$n_rows` country-day observations.

```{r show-country-table}
countries <- daily_panel %>% 
  distinct(who_region, who_region_long, country_name) %>% 
  arrange(who_region) %>% 
  mutate(who = paste0(who_region_long, " (", who_region, ")")) %>% 
  group_by(who) %>% 
  nest() %>% 
  mutate(nice_matrix = map(data, ~matrix_from_vector(.$country_name, ncol = 6))) %>% 
  mutate(matrix_rows = map_int(nice_matrix, ~nrow(.)))

row_groups <- countries$matrix_rows %>% 
  set_names(countries$who)

reduce(countries$nice_matrix, rbind) %>% 
  kbl(table.attr = "class=\'table-sm\'") %>% 
  kable_styling(bootstrap_options = c("basic"), 
                full_width = FALSE) %>% 
  pack_rows(index = row_groups, indent = FALSE, 
            label_row_css = 'background-color: rgb(2, 184, 117); color: #ffffff; text-align: center;')
```

### General overview of full data

Here's an overview of the complete panel data:

```{r glimpse-everything}
#| code-fold: false
glimpse(daily_panel)
```

## ICCPR derogations

Our data is based on reading the texts of ICCPR derogations at the UN Treaty Collection site at <https://treaties.un.org/>. We create these variables:

- **ICCPR derogation filed** (`iccpr_derogation_filed`): Date that derogation notice was filed/received by the UN. This is a dichotomous variable that is `TRUE` on the day the country files the notice and `FALSE` otherwise.
- **Derogation start** (`derogation_start`): This is the date noted within the derogation text of when the emergency provisions and suspension of rights starts. This is a dichotomous variable that is `TRUE` on the day the country starts its derogation and `FALSE` otherwise.
- **Derogation in effect** (`derogation_ineffect`): This is a dichotomous variable that is `TRUE` if the country is currently derogating and `FALSE` if not.
- **Derogation end** (`derogation_end`): This is the date noted within the derogation text of when emergency provisions stop and rights observances resume. This is a dichotomous variable that is `TRUE` on the day the country finishes its derogation and `FALSE` otherwise.

:::{.callout-note}
In many cases states filed derogations after the derogations actually started, so filed dates are often after the derogation start dates. Also note that states often updated the timing of derogations to extend, so there may be multiple filing dates within one derogation period.
:::

Here's an overview of these variables:

```{r glimpse-iccpr}
#| code-fold: false
daily_panel %>%
  select(country_name, iso3, day, iccpr_derogation_filed,
         derogation_start, derogation_ineffect, derogation_end) %>% 
  glimpse()
```

And here's an extract of the data from June 11, 2020, the day that Argentina filed a derogation notice:

```{r show-iccpr}
daily_panel %>%
  filter(day == ymd("2020-06-11")) %>%
  select(country_name, iso3, day, iccpr_derogation_filed,
         derogation_start, derogation_ineffect, derogation_end) %>% 
  datatable()
```

## WHO COVID case counts

We collect daily case counts from the [WHO's COVID-19 Dashboard](https://covid19.who.int/data). We use these variables:

- **WHO region** (`who_region` and `who_region_long`): One of six WHO regional offices:
  - Regional Office for Africa (AFRO)
  - Regional Office for the Americas (AMRO)
  - Regional Office for the Eastern Mediterranean (EMRO)
  - Regional Office for Europe (EURO)
  - Regional Office for South-East Asia (SEARO)
  - Regional Office for the Western Pacific (WPRO)
- **New cases** (`new_cases`): New confirmed cases of COVID-19. Calculated by subtracting previous cumulative case count from current cumulative cases count.
- **Cumulative cases** (`cumulative_cases`): Cumulative confirmed cases of COVID-19 reported to the WHO to date.
- **New deaths** (`new_deaths`): New confirmed COVID-19 deaths. Calculated by subtracting previous cumulative deaths from current cumulative deaths.
- **Cumulative deaths** (`cumulative_deaths`): Cumulative confirmed COVID-19 deaths reported to the WHO to date.

We omit all "other" observations not assigned to any country (like cruise ships), and remove all non-states that don't appear in the UN's ICCPR list or in V-Dem, including:

```{r show-nonstate-table}
nonstates <- c("Falkland Islands", "Fore Islands", "Guernsey", "Guam", 
               "Isle of Man ", "Jersey", "Saint Pierre and Miquelon", 
               "Pitcairn Islands", "Puerto Rico", "Reunion Island", "Tokelau", 
               "Holy See", "Wallis and Futuna", "Bonaire", "Sint Eustatius", 
               "Saba", "Mayotte", "Gibraltar ", "Greenland", "Guadeloupe")

matrix_from_vector(nonstates, ncol = 5) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("basic"), 
                full_width = FALSE) %>% 
  pack_rows("Omitted non-states", 1, 4, indent = FALSE, 
            label_row_css = 'background-color: rgb(184, 71, 67); color: #ffffff; text-align: center;')
```

Here's an overview of these variables:

```{r glimpse-who}
#| code-fold: false
daily_panel %>%
  select(country_name, iso3, day, who_region, who_region_long,
         new_cases, cumulative_cases, new_deaths, cumulative_deaths) %>% 
  glimpse()
```

## Oxford government responses

To measure the severity of government responses to COVID-19, we use data from the [Oxford COVID-19 Government Response Tracker](https://www.bsg.ox.ac.uk/research/research-projects/covid-19-government-response-tracker). We use the project's overall stringency index, as well as several measures of economic policies and containment and closure policies:

- **Stringency index** (`stringency_index`): TODO
- **Cancel public events** (`c3_cancel_public_events`): Indicator of whether the country cancelled public events.
  - **0**: no measures; **1**: recommend cancelling; **2**: require cancelling; **Blank**: no data
- **Limits on gatherings** (`c4_restrictions_on_gaterings`): Indicator of whether the country limited public gatherings.
  - **0**: no restrictions; **1**: restrictions on very large gatherings (the limit is above 1000 people); **2**: restrictions on gatherings between 101-1000 people; **3**: restrictions on gatherings between 11-100 people; **4**: restrictions on gatherings of 10 people or less; **Blank**: no data
- **Close public transport** (`c5_close_public_transport`): Indicator of whether the country closed public transport.
  - **0**: no measures; **1**: recommend closing (or significantly reduce volume/route/means of transport available); **2**: require closing (or prohibit most citizens from using it); **Blank**: no data
- **Stay at home requirements** (`c6_stay_at_home_requirements`): Indicator of whether the country imposed shelter-in-place requirements.
  - **0**: no measures; **1**: recommend not leaving house; **2**: require not leaving house with exceptions for daily exercise, grocery shopping, and 'essential' trips; **3**: require not leaving house with minimal exceptions (eg allowed to leave once a week, or only one person can leave at a time, etc); **Blank**: no data
- **Restrictions on internal movement** (`c7_movement_restrictions_intern`): Indicator of whether the country imposed restrictions on internal movement between cities or regions.
  - **0**: no measures; **1**: recommend not to travel between regions/cities; **2**: internal movement restrictions in place; **Blank**: no data
- **International travel controls** (`c8_international_travel_control`): Indicator of whether the country imposed restrictions on international travel (for foreign travellers, not citizens)
  - **0**: no restrictions; **1**: screening arrivals; **2**: quarantine arrivals from some or all regions; **3**: ban arrivals from some regions; **4**: ban on all regions or total border closure; **Blank**: no data
- **Income support** (`e1_income_support`): Indicator of whether the country provided direct cash payments for those unable to work.
  - **0**: no income support; **1**: government is replacing less than 50% of lost salary (or if a flat sum, it is less than 50% median salary); **2**: government is replacing 50% or more of lost salary (or if a flat sum, it is greater than 50% median salary); **Blank**: no data
- **Debt relief** (`e2_debt_relief`): Indicator of whether the country froze financial obligations for households (e.g. loan freezes, eviction bans, etc.)
  - **0**: no debt/contract relief; **1**: narrow relief, specific to one kind of contract; **2**: broad debt/contract relief; **Blank**: no data

Here's an overview of these variables:

```{r glimpse-oxford}
#| code-fold: false
daily_panel %>% 
  select(country_name, iso3, day, stringency_index, c3_cancel_public_events,
         c4_restrictions_on_gaterings, c5_close_public_transport,
         c6_stay_at_home_requirements, c7_movement_restrictions_intern,
         c8_international_travel_control, e1_income_support, e2_debt_relief) %>% 
  glimpse()
```

And here's an extract of the data from June 11, 2020:

```{r show-oxford}
daily_panel %>%
  filter(day == ymd("2020-06-11")) %>%
  select(country_name, iso3, day, stringency_index, c3_cancel_public_events,
         c4_restrictions_on_gaterings, c5_close_public_transport,
         c6_stay_at_home_requirements, c7_movement_restrictions_intern,
         c8_international_travel_control, e1_income_support, e2_debt_relief) %>% 
  datatable()
```

## V-Dem

TODO