### 2022-07-24 analysis/analysis_h1.qmd
Add hypotheses and other placeholders

---
title: "1: Derogations and human rights"
format:
  html:
    code-fold: true
---