### 2022-07-24 analysis/analysis_h3.qmd

---
title: "3: Civil society, derogations, and economic policies"
format:
  html:
    code-fold: true
---