### 2022-07-25 analysis/data-final.qmd
Use file-based targets + add data to website

---
title: "Download final data"
format:
  html:
    code-tools: false
---

<table class="table">
  <thead>
    <tr>
      <th scope="col">Program</th>
      <th scope="col">File</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>CSV file for any program</td>
      <td><a class="btn btn-primary" target="_blank" href="data/daily_panel.csv">{{< fa table >}} &ensp;<span class="font-monospace">daily_panel.csv</span></a></td>
    </tr>
    <tr>
      <td><span class="font-monospace">.rds</span> file for R<br><span class="small">(load with `df <- readRDS("daily_panel.rds")`)</small></td>
      <td><a class="btn btn-primary" target="_blank" href="data/daily_panel.rds">{{< fa brands r-project >}} &ensp;<span class="font-monospace">daily_panel.rds</span></a></td>
    </tr>
    <tr>
      <td><span class="font-monospace">.dta</span> file for Stata<br><span class="small">(load with `use "daily_panel.dta"`)</small></td>
      <td><a class="btn btn-primary" target="_blank" href="data/daily_panel.dta">{{< fa table-cells >}} &ensp;<span class="font-monospace">daily_panel.dta</span></a></td>
    </tr>
  </tbody>
</table>