### 2023-02-14 analysis/presentations/2022-11-04_peace-science.qmd

---
title: "Pandemic Pass"
subtitle: "Treaty Derogations and<br>Human Rights Practices During COVID-19"
author:
  - name: Suparna Chaudhry
    affiliation: Lewis and Clark College
  - name: Audrey Comstock
    affiliation: Arizona State University
  - name: Andrew Heiss
    affiliation: Georgia State University
date: "November 4, 2022"
date-format: long
output: revealjs
format:
  revealjs: 
    slide-number: true
    chalkboard:
      buttons: true
    preview-links: auto
    footer: "Chaudhry, Comstock, and Heiss&emsp;|&emsp;Pandemic Pass"
    # include-before-body: testing.html
    # revealjs-plugins:
    #   - simplemenu
    theme: [simple, custom.scss]
    title-slide-attributes:
      data-background-color: "#c06636"
# resources:
  # - slides.pdf
---

```{r include=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(scales)
library(marginaleffects)
library(patchwork)
library(sf)

knitr::opts_chunk$set(fig.retina = 3, collapse = TRUE, echo = FALSE)
options(digits = 3, width = 75)

# Load targets 
withr::with_dir(here::here(), {
  tar_load(year_week_lookup)
  tar_load(civicus_map_data)

  # Models
  tar_load(m_prelim_derog)
  tar_load(m_h1)
  tar_load(m_h2)
  tar_load(m_h3)

  # Plotting functions
  invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
})

clrs_lt <- colorspace::lighten(clrs, 0.9)
```

```{r make-plots, include=FALSE, cache=TRUE}
preliminary_models_plots <- m_prelim_derog %>% 
  filter(y != "c8_intl_travel_bin") %>% 
  mutate(pred_plot = map2(model, nice, ~{
    datagrid(model = .x,
             year_week_num = 1:69,
             derogation_ineffect = 0:1) %>% 
      add_epred_draws(.x) %>% 
      left_join(year_week_lookup, by = "year_week_num") %>% 
      mutate(derogation_ineffect = factor(derogation_ineffect, 
                                          levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
      ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_fill_manual(values = c(clrs[1], clrs[8])) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      scale_y_continuous(labels = label_percent()) +
      labs(x = NULL, y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  }))

h1_models_plots <- m_h1 %>% 
  mutate(pred_plot = pmap(list(model, nice, family), ~{
    if (..3 == "cumulative") {
      datagrid(model = ..1,
               year_week_num = 1:69,
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        left_join(year_week_lookup, by = "year_week_num") %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No derogation", "Derogation in effect"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = year_week_day, y = .epred, color = .category)) +
        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        scale_y_continuous(labels = label_percent()) +
        labs(x = NULL, y = "Predicted probability", 
             color = "Severity", fill = "Severity",
             title = ..2) +
        facet_wrap(vars(derogation_ineffect)) +
        theme_pandem()
    } else {
      datagrid(model = ..1,
               year_week_num = 1:69,
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        left_join(year_week_lookup, by = "year_week_num") %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No", "Yes"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = year_week_day, y = .epred, color = derogation_ineffect)) +
        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
        scale_color_manual(values = c(clrs[1], clrs[8])) +
        scale_fill_manual(values = c(clrs[1], clrs[8])) +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        scale_y_continuous(labels = label_percent()) +
        labs(x = NULL, y = "Predicted probability", 
             color = "Derogation in effect", fill = "Derogation in effect",
             title = ..2) +
        theme_pandem()
    }
  }))

h2_models_plots <- m_h2 %>% 
  mutate(pred_plot = pmap(list(model, nice, family), ~{
    if (..3 == "cumulative") {
      datagrid(model = ..1,
               v2csreprss = seq(-2.5, 3, by = 0.1),
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No derogation", "Derogation in effect"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = v2csreprss, y = .epred, color = .category)) +
        stat_lineribbon(aes(fill = .category), alpha = 0.3) +
        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_fill_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_y_continuous(labels = label_percent()) +
        labs(x = "Civil society repression", y = "Predicted probability", 
             color = "Severity", fill = "Severity",
             title = ..2) +
        facet_wrap(vars(derogation_ineffect)) +
        theme_pandem()
    } else {
      datagrid(model = ..1,
               v2csreprss = seq(-2.5, 3, by = 0.1),
               derogation_ineffect = 0:1) %>% 
        add_epred_draws(.x) %>% 
        mutate(derogation_ineffect = factor(derogation_ineffect, 
                                            levels = 0:1,
                                            labels = c("No", "Yes"),
                                            ordered = TRUE)) %>% 
        ggplot(aes(x = v2csreprss, y = .epred, color = derogation_ineffect)) +
        stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
        scale_color_manual(values = c(clrs[1], clrs[8])) +
        scale_fill_manual(values = c(clrs[1], clrs[8])) +
        scale_y_continuous(labels = label_percent()) +
        labs(x = "Civil society repression", y = "Predicted probability", 
             color = "Derogation in effect", fill = "Derogation in effect",
             title = ..2) +
        theme_pandem()
    }
  }))

h3_models_plots_cs <- m_h3 %>% 
  mutate(pred_plot = map2(model, nice, ~{
    datagrid(model = .x,
             v2xcs_ccsi = seq(0, 1, by = 0.01),
             derogation_ineffect = 0:1) %>% 
      add_epred_draws(.x) %>% 
      mutate(derogation_ineffect = factor(derogation_ineffect, 
                                          levels = 0:1,
                                      labels = c("No", "Yes"),
                                      ordered = TRUE)) %>% 
      ggplot(aes(x = v2xcs_ccsi, y = .epred, color = derogation_ineffect)) +
      stat_lineribbon(aes(fill = derogation_ineffect), alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_fill_manual(values = c(clrs[1], clrs[8])) +
      scale_y_continuous(labels = label_percent()) +
      labs(x = "Core civil society index", y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  }))
```

# Question {background-color="`r clrs[3]`" .white}

##

![](img/covid-headlines@3x.png){fig-align="center" width="100%"}

## Emergencies and international law {.clrs-3}

### Derogations

Provisions in international law that authorize states to temporarily suspend treaty obligations

. . .

Many states do not formally derogate and just implement emergency measures

## ICCPR derogations during COVID-19 {.clrs-3}

::: {.smaller}

| Type of response                                              |   N |
|:--------------------------------------------------------------|----:|
| Derogated as intended                                         |  13 |
| Derogated; measures not temporary, proportional, or necessary |   5 |
| Emergency declared; no formal derogation                      |  74 |
| No emergency; no formal derogation                            |  28 |
| Has not ratified ICCPR                                        |   8 |

:::

## Research questions {.clrs-3}

::: {.incremental}
- Were derogating states less likely to abuse human rights as they implemented emergency measures for COVID-19?
- Can civil society organizations (CSOs) mitigate the adverse effects of these emergency measures?
:::

## Preview of findings {.clrs-3}

Derogating states are less likely to abuse emergency measures, have discriminatory policies, violate non-derogable rights, and impose measures without time limits

. . .

Permissive civil society environment increases likelihood of debt relief policies

. . .

Civil society repression increases the probability of abusive enforcement and discriminatory policies; derogations mitigate that probability

## Significance {.clrs-3}

**International law and human rights**

International treaty provisions designed to safeguard rights may end up worsening conditions when there is a lack of review from the international community

. . .

**Democratic erosion and human rights**

Emergencies are important for understanding the processes of autocratization and democratic backsliding

# Theory and hypotheses {background-color="`r clrs[5]`" .white}

## UN requirements for public health emergency measures {.clrs-5}

Suspensions of rights must be:

- Legal
- Necessary
- Proportional
- Non-discriminatory

## Derogations during COVID-19 {.clrs-5}

```{r prelim-plots}
#| echo: false
#| fig-width: 9
#| fig-height: 5
#| out-width: 100%
#| fig-align: center
wrap_plots(preliminary_models_plots$pred_plot)
```

## Pandemic as excuse {.clrs-5}

If states implement emergency measures without regard for international law, the implementation may be problematic and the process to restore rights may not be followed

- Restrictions on travel or movement by minority or opposition groups
- Limits on public assembly and protests
- Deployment of surveillance systems that impede privacy

## H~1~: Derogations & human rights {.clrs-5}

If states formally derogate from international treaty provisions, then they are less likely to:

a. Use discriminatory policy measures
b. Violate non-derogable rights
c. Impose measures with no time limits

## Civil society and human rights {.clrs-5}

Civil society helps mitigate the impact of crises, emergencies, and natural disasters

:::: {.columns}

::: {.column width="50%" .smallest}
During COVID-19, civil society:

- Broadened range of voices in policymaking process
- Pursued civic initiatives to deal with impact of economic shutdown, product shortages, and spikes in unemployment
- Assisted with public health initiatives
:::

::: {.column width="50%" .smaller}
![](img/image5.png){fig-align="center" width="100%"}
:::

::::

## H~2~: Civil society & supportive policies {.clrs-5}

If civil society organizations are involved in the policy making process, then states are more likely to have supportive economic policies

## Pandemic as excuse {.clrs-5}

Some states used emergency measures as an excuse to limit democratic participation, dissent, and challenges to economic and political rule

![](img/covid-excuse@3x.png){fig-align="center" width="100%"}

## Civil society repression {.clrs-5}

```{r civicus-map}
#| echo: false
#| fig-width: 6.5
#| fig-height: 3.1
#| out-width: 100%
#| fig-align: center
#| fig-cap: "2020 CIVICUS Monitor civic space ratings"
ggplot() +
  geom_sf(data = civicus_map_data, aes(fill = fct_rev(category)), size = 0.15, color = "black") +
  coord_sf(crs = st_crs("ESRI:54030"), datum = NA) +  # Robinson
  scale_fill_manual(values = clrs[c(8, 7, 4, 2, 1)],
                    na.translate = FALSE, name = "Civic space") +
  theme_void(base_family = "Noto Sans") +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.7, "lines"))
```

## H~3~: Civil society & human rights {.clrs-5}

If states are repressive towards civil society organizations, then they are more likely to:

a. Violate non-derogable rights
b. Use discriminatory policy measures
c. Engage in abusive enforcement of emergency measures
d. Impose limits on media freedom

# Data and results {background-color="`r clrs[1]`" .white}

## Data {.clrs-1}

- Hand-coded derogations from the ICCPR
- Oxford's COVID-19 Government Response Tracker
- V-Dem's Pandemic Backsliding Project
- WHO's case and death counts

. . .

Daily data for 139 countries from March 11, 2020 to June 30, 2021

## Modeling approach {.clrs-1}

:::: {.columns}

::: {.column width="50%" .smallest}

**Outcomes**

- Discriminatory policies (H~1~, H~3~)
- Violation of non-derogable rights (H~1~, H~3~)
- Abusive enforcement (H~1~, H~3~)
- Measures without time limits (H~1~)
- Limiting media (H~3~)
- Income support (H~2~)
- Debt relief (H~2~)

:::

::: {.column width="50%" .smallest}
**Explanatory variables**

- Derogation in effect
- Civil society repression
- Core civil society index

**Other controls**

- COVID case counts
- COVID death counts
- Rule of law index
- Civil liberties index

:::

::::

## Modeling approach {.clrs-1}

Bayesian multilevel models with country random effects and weekly time trend

. . .

*(Logistic models for binary outcomes; ordered logistic models for outcomes with multiple categories)*

. . .

**Open to other ideas too!**

## H~1~: Derogations and human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smaller}
Substantial 2–3 percentage point difference in probability of violating non-derogable rights (P[∆ < 0] = 0.964)
:::

::: {.column width="70%"}
```{r include=FALSE, eval=FALSE}
qwer <- m_h1 %>% 
  filter(y == "pandem_ndrights") %>% 
  pull(model) %>% pluck(1) %>% 
  marginaleffects(newdata = datagrid(year_week_num = c(1, 69)),
                  variables = "derogation_ineffect",
                  type = "response") %>% 
  posteriordraws() %>% 
  left_join(year_week_lookup, by = "year_week_num")

qwer %>% 
  group_by(year_week_day) %>% 
  summarize(avg_effect = mean(draw), 
            prop_less_0 = sum(draw < 0) / n())

ggplot(qwer, aes(x = year_week_day, y = draw * 100)) +
  stat_lineribbon()
```

```{r h1-nonderog}
#| fig-width: 4.5
#| fig-height: 3
#| out-width: 100%
#| fig-align: center
h1_models_plots %>% 
  filter(y == "pandem_ndrights") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Violation of non-derogable rights")
```
:::

::::

## H~1~: Derogations & human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smaller}
Substantial ≈22 percentage point difference in probability of having time-limited measures (P[∆ ≠ 0] = 1)
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h1 %>% 
  filter(y == "pandem_nolimit") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(year_week_num = c(1)),
                  variables = "derogation_ineffect",
                  type = "response")
```

```{r h1-timelimit, message=FALSE}
#| fig-width: 5
#| fig-height: 3.65
#| out-width: 100%
#| fig-align: center
h1_models_plots %>% 
  filter(y == "pandem_nolimit") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Emergency measures with no time limits") +
  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y",
               guide = guide_axis(n.dodge = 2))
```
:::

::::

## H~1~: Derogations & human rights {.clrs-1}

### Null results

No derogation-based difference in predicted probability of having discriminatory policies or abusive enforcement.

## H~2~: Civil society & supportive policies {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smaller}
Both derogation status and civil society strength substantially increase the probability of a state providing COVID-related debt relief
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h3 %>% 
  filter(y == "e2_debt_relief_bin") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(v2xcs_ccsi = c(0, 0.5, 0.75, 1)),
              variables = "derogation_ineffect",
              type = "response")
```

```{r h3-incomesupport, message=FALSE}
#| fig-width: 4.5
#| fig-height: 3
#| out-width: 100%
#| fig-align: center
h3_models_plots_cs %>% 
  filter(y == "e2_debt_relief_bin") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Debt relief policies")
```
:::

::::

## H~3~: Civil society & human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smallest}
Substantial difference in probability of implementing discriminatory policies across all levels of civil society repression (P[∆&nbsp;≠&nbsp;0] = 1) for states that derogated

Better civil society associated with less discriminatory policy in derogating states
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h2 %>% 
  filter(y == "pandem_discrim") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(v2csreprss = c(-2, 0, 2)),
                  variables = "derogation_ineffect",
                  type = "response")
```

```{r h2-discrim}
#| fig-width: 5
#| fig-height: 3.65
#| out-width: 100%
#| fig-align: center
h2_models_plots %>% 
  filter(y == "pandem_discrim") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Discriminatory policies")
```
:::

::::

## H~3~: Civil society & human rights {.clrs-1}

:::: {.columns}

::: {.column width="30%" .smallest}
Substantial difference in probability of abusive enforcement of emergency measures for countries with worse civil society repression in derogating states
:::

::: {.column width="70%"}
```{r eval=FALSE, include=FALSE}
m_h2 %>% 
  filter(y == "pandem_abusive") %>% 
  pull(model) %>% pluck(1) %>% 
  comparisons(newdata = datagrid(v2csreprss = c(-2, 0, 2)),
              variables = "derogation_ineffect",
              type = "response")
```

```{r h2-abusive}
#| fig-width: 5
#| fig-height: 3.65
#| out-width: 100%
#| fig-align: center
h2_models_plots %>% 
  filter(y == "pandem_abusive") %>% 
  pull(pred_plot) %>% pluck(1) +
  labs(title = "Abusive enforcement")
```
:::

::::

## H~3~: Civil society & human rights {.clrs-1}

### Null results

No derogation-based difference in predicted probability of violating non-derogable rights or limiting media across range of civil society repression

# Discussion {background-color="`r clrs[8]`" .white}

## Summary {.clrs-8}

Derogating states are less likely to abuse emergency measures, have discriminatory policies, violate non-derogable rights, and impose measures without time limits

. . .

Civil society repression increases the probability of abusive enforcement and discriminatory policies; derogations mitigate that probability

. . .

Permissive civil society environment increases likelihood of debt relief policies

## Broader implications {.clrs-8}

Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties

. . .

Some states use derogations as intended; others use them as an excuse to violate rights

. . .

Understanding how states treat derogations is important for understanding declining respect for human rights and democratic backsliding