### 2023-03-08 analysis/manuscript.qmd
Add special manuscript iframe page

---
title: Manuscript
toc: false
page-layout: full
---

<p class="text-center"><a class="btn btn-success" href="/manuscript/output/manuscript.html" target="manuscript">{{< fa file-lines >}} &ensp;HTML</a> <a class="btn btn-success" href="/manuscript/output/manuscript.html" target="_blank">{{< fa arrow-up-right-from-square >}} &ensp;HTML (new window)</a> <a class="btn btn-success" href="/manuscript/output/manuscript.pdf" target="manuscript">{{< fa file-pdf >}} &ensp;PDF</a> <a class="btn btn-success" target="_blank" href="/manuscript/output/manuscript.pdf">{{< fa file-arrow-down >}} &ensp;PDF (download)</a></p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="manuscript"></iframe>