### 2023-03-08 index.qmd
Use newer abstract

---
title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
format:
  html:
    code-tools: false
sidebar: false
---

- Suparna Chaudhry (Lewis & Clark College)
- Audrey L. Comstock (Arizona State University)
- Andrew Heiss (Georgia State University)

In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations. Did states actually end up complying with their pandemic-era derogations? Did derogating and non-derogating states implement emergency measures differently, and if so, how? Using data from the Varieties of Democracy PanDem dataset and the Oxford Covid-19 Government Response Tracker, we find that states which derogated from their international human rights obligations imposed emergency measures that were temporary and non-discriminatory. On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights. Our results support the role that flexibility mechanisms such as derogations play in international law.