### 2023-03-11 presentations/2023-03-17_isa/index.qmd
Add placeholder page for ISA

---
title: "International Studies Association"
date: 2023-03-17
toc: false
resources: slides
abstract: |
  Annual conference of the International Studies Association (ISA); Montréal, Canada
abstract-title: "Location"
published-title: "Presented"
# image: first-slide.png
page-layout: full
---

Slides here.