### 2023-03-11 presentations/2023-03-17_isa/index.qmd
Embed Google slides

---
title: "International Studies Association"
date: 2023-03-17
toc: false
resources: slides
abstract: |
  Annual conference of the International Studies Association (ISA); Montréal, Canada
abstract-title: "Location"
published-title: "Presented"
image: first-slide.png
page-layout: full
---

<div class="ratio ratio-16x9">
<iframe class="slide-deck" src="https://docs.google.com/presentation/d/e/2PACX-1vQl8N_xTO7QyJNFuEvJXrUDV77_DLbgnjZ4X6WFx0q3ScXPEa6Y-uvUanzHs2ds5Gt8E-5Px0Biqa3l/embed?start=false&loop=false&delayms=3000" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
</div>