### 2023-03-11 analysis/hypotheses.qmd
Remove questions and methods section

---
title: "Hypotheses"
format:
  html:
    code-tools: false
---

We test two sets of hypotheses:

## H~1~: Formal derogations and pandemic-related human rights violations

- **H~1a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
- **H~1b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
- **H~1c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.

## H~2~: Civil society repression and pandemic-related human rights violations

- **H~2a~**: If states are repressive towards civil society organizations, then they are more likely to violate non-derogable rights.
- **H~2b~**: If states are repressive towards civil society organizations, then they are more likely to use discriminatory policy measures during the pandemic.
- **H~2c~**: If states are repressive towards civil society organizations, then they are more likely to engage in abusive enforcement of measures instated to deal with the pandemic.
- **H~2d~**: If states are repressive towards civil society organizations, then they are more likely to impose limits on media freedom during the pandemic.