### 2023-03-11 analysis/manuscript.qmd
Nicer button colors

---
title: Manuscript
toc: false
page-layout: full
---

```{r include=FALSE}
# Add the manuscript target as a dependency to the overall website
targets::tar_load(c(manuscript_nice, manuscript_manuscripty))
```

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="manuscript">
  <i class="fa-solid fa-file-lines"></i> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="_blank">
  <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/manuscript.pdf" target="manuscript">
  <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
  </a>

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>

  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/manuscript-ms.pdf" target="manuscript">
  <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
  </a>

  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/manuscript-ms.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>
  </a>
</p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="manuscript"></iframe>
```