### 2023-04-20 analysis/hypotheses.qmd

---
title: "Hypotheses"
format:
  html:
    code-tools: false
---

We test two sets of hypotheses:

## H~1~: Formal derogations and pandemic-related emergency policies

- **H~1a~**: If states issued COVID-19 derogations, then we expect that they will be more likely to impose relevant emergency polices such as canceling public events, restricting gatherings, closing public transit, and restricting domestic and international movement.
- **H~1b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose related emergency policies.

## H~2~: Formal derogations and pandemic-related human rights violations

- **H~2a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
- **H~2b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
- **H~2c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.