### 2023-05-05 manuscript/manuscript.qmd
woohoo dynamic in-text numbers

---
title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
short-title: "Pandemic Pass?"
subtitle: "(Research note)"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  orcid: 0000-0003-3506-7590
  url: https://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Audrey L. Comstock
  affiliation: Arizona State University
  orcid: 0000-0003-1772-5350
  url: https://audreylcomstock.weebly.com/
  email: audrey.comstock@asu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  orcid: 0000-0002-3948-3914
  url: https://www.andrewheiss.com
  email: aheiss@gsu.edu
abstract: |
  This research note asks whether states issuing pandemic-era human rights treaty derogations implemented emergency provisions as intended or used them to abuse human rights during a time of crisis. In an effort to combat the COVID-19 pandemic, many countries declared states of emergency and derogated (temporarily suspended) from their international human rights treaty obligations. Using data from the Varieties of Democracy PanDem dataset and the Oxford COVID-19 Government Response Tracker, we find that states that derogated from their international human rights obligations imposed emergency measures that were temporary and did not violate non-derogable rights. On the other hand, states that did not derogate were more likely impose discriminatory measures, enact emergency measures without time limits and violate non-derogable rights. Our results support the role that flexibility mechanisms such as derogations play in international law and show that states are being sincere about their intentions and not, generally, using these mechanisms to cover abusive behavior.
date: May 2, 2023
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
keywords:
  - human rights
  - treaty derogations
  - COVID-19
  - emergency policies
bibliography: bib.bib
link-citations: true
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
                      warning = FALSE, message = FALSE)

knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE)

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r libraries-data}
library(tidyverse)
library(scales)
library(sf)
library(lubridate)
library(tidybayes)
library(marginaleffects)
library(patchwork)
library(colorspace)
library(modelsummary)
library(kableExtra)
library(gt)
library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Generated via random.org
set.seed(8511)

# Load targets 
weekly_panel <- tar_read(weekly_panel) %>% 
  mutate(year_week_day = ymd(year_week_day))

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))
tar_load(c(derogation_map_data, derogation_count))

# Plotting and modelsummary functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))
```

```{r make-var-lists-and-tbls}
derog_detail <- derogation_count %>% 
  filter(derogations > 0) %>% 
  summarize(countries = n(),
            total = sum(derogations)) %>% 
  as.list()

map_data_nice <- derogation_count %>% 
  filter(derogations > 0) %>% 
  arrange(desc(derogations)) %>% 
  mutate(Country = countrycode::countrycode(iso3, "iso3c", "country.name")) %>% 
  select(Country, Derogations = derogations)

fmt_p_inline <- function(x, direction) {
  x <- round(x, 2)

  if (direction == "gt") {
    out <- glue::glue(r"[$p(\Delta > 0) = {x}$]")
  } else {
    out <- glue::glue(r"[$p(\Delta < 0) = {x}$]")
  }

  return(out)
}

policies_pred <- policies_plot_data %>% 
  unnest(pred_data) %>% 
  filter(.width == 0.95) %>% 
  select(nice, year_week_day, derogation_ineffect, .epred, .lower, .upper) %>% 
  group_by(nice, derogation_ineffect) %>% 
  mutate(rank = dense_rank(.epred)) %>% 
  filter(rank == 1 | rank == max(rank)) %>% 
  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
  group_by(rank, derogation_ineffect) %>% 
  mutate(outcome = janitor::make_clean_names(nice)) %>%
  rename(draw = .epred, min = .lower, max = .upper) %>% 
  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0)))) %>% 
  split(~outcome + derogation_ineffect + rank)

policies_mfx <- policies_plot_data %>% 
  unnest(mfx_data) %>% 
  filter(.width == 0.95) %>% 
  select(nice, year_week_day, draw, .lower, .upper, p_gt_0, .width) %>% 
  group_by(nice) %>% 
  mutate(rank = dense_rank(draw)) %>% 
  filter(rank == 1 | rank == max(rank)) %>% 
  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
  group_by(rank) %>% 
  mutate(outcome = janitor::make_clean_names(nice)) %>%
  rename(min = .lower, max = .upper) %>% 
  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0))),
         p_lt_0 = p_gt_0 - 1) %>% 
  mutate(p_gt = fmt_p_inline(p_gt_0, "gt"),
         p_lt = fmt_p_inline(p_lt_0, "lt")) %>% 
  split(~outcome + rank)
```

What is the impact of international legal emergency provisions on human rights? Combating a deadly virus like COVID-19 required extraordinary public health measures. In an effort to fight the COVID-19 pandemic, governments across the world introduced restrictions on the freedom of movement, association, and assembly. While countries were legally able to introduce these policies within their borders, the restrictions raised important questions about states’ commitments to international human rights law. Some human rights treaties recognize states’ ability to use emergency measures during times of crises, and contain "escape clauses" that safeguard the permanent erosion of human rights during and following emergencies. The temporary suspension of state compliance from international treaty obligations occurs through a legal action called derogation. However, the potential abuse of emergency measures during derogations raises concerns about their impact on human rights. Under emergency decrees, states often impose restrictions that are temporary in theory but difficult to roll back or contain in practice  [@Stasavage2020; @LuhrmannRooney2021]. Did states use pandemic era derogations how they were intended or were they used as a "pandemic pass" to justify violating human rights?

Given recent medical advancements such as the distribution of vaccines globally and declining concern about COVID-19 severity, the Biden administration announced that it would end the public health emergency in the US on May 11, 2023. Doing so lifts measures instituted at the beginning of the pandemic—a move that has occurred in many other countries as well. As such, it is a worthwhile time to assess important descriptive and analytical questions about states’ respect for international law and human rights during this emergency period. This research note examines what derogating meant for compliance with international law during the COVID-19 pandemic. Specifically, we examine whether states actually derogated as per international legal principles and investigate the broader concern about whether states used derogations to exploit emergency provisions and to abuse human rights during COVID-19. These questions are important for the study of international law, compliance, and human rights during times of crises.

Using data from the Varieties of Democracy PanDem dataset, the Oxford COVID-19 Government Response Tracker, and United Nations treaty derogation data, we first assess the variation among derogating and non-derogating states on the implementation of emergency measures. We then analyze the effect of derogating (or not) on human rights practices during the pandemic. We find that states that derogated and specified time limits to these emergency provisions generally stuck to the time limits and adhered to standards required during derogation by avoiding discriminatory policies and not violating non-derogable rights (the right to life, freedom from torture, and freedom from slavery). On the other hand, states that did not derogate from their international human rights obligations and still imposed emergency measures were more likely to impose discriminatory policies, violate non-derogable rights, and impose emergency measures without time limits.

This research note makes two contributions. First, while there is a significant literature on compliance with international law, especially human rights treaties [e.g. @Simmons2009; @Comstock2021], and the flexibility tools such as reservations, understandings, and declarations (RUDs) that states invoke when they join or withdraw from such treaties [@Neumayer:2007; @Helfer:2002; @Helfer:2005], we know little about states' compliance with flexibility mechanisms such as derogations. Compliance with derogations may not necessarily be the same as compliance with the treaty and this research note provides a systematic understanding of adherence to derogations. When states comply with derogations, they are actually limiting their obligations to the human rights treaty overall. As such, we should in fact expect a reduction in compliance with the overall human rights treaty during that time. We explore how the potential decline in compliance may actually be a good indication that states are using treaty design as intended.

The pandemic further provides unique insight into derogation behavior because numerous states derogated under the same global conditions of a public health emergency, rather than domestic conflict or natural disaster, typical of treaty derogations, making it easier to compare state behavior during these derogations cross nationally. In making this comparison, this research note makes an important descriptive contribution showing how the implementation of policy measures differed across derogating and non-derogating states during the pandemic.

Second, there is a robust literature on compliance with international law and how it can lead to improvements in domestic human rights conditions [@Simmons2009; @Merry:2009]. While derogations are intended to make sure government responses are temporary, proportional and non-discriminatory, some legal scholars warn of the possibility that some treaty actions removing legal obligations may actually contribute to more violations and abuse [e.g. @Lebret:2020; @helfer2021], as part of a broader trend of targeting groups during times of crises [@Grasse2021]. We find that overall, derogating states during the pandemic did not increase abuse and violations. This finding speaks to international legal compliance scholarship and supports the idea that states are being sincere about their intentions and not, generally, using actions like derogation to cover abusive behavior.

The rest of the paper proceeds as follows. First, we explore the purpose of derogations in international law, and then specifically during the pandemic. In doing so, we focus on derogations to the International Covenant on Civil and Political Rights (ICCPR)—arguably the most significant and extensive treaty in the international human rights regime. Following this, we introduce our expectations regarding how formally derogating from the treaty (or not) may impact a state's human rights practices during the pandemic. We conclude with the implications of our research note for the study of international law during times of crises.

# Derogations and international law

Derogations in international law authorize states to temporarily suspend their international treaty commitments during times of crises, providing them with flexibility to respond to events. Times of crises can include wars, natural disasters, internal political unrest such as terrorism, and public health emergencies. Derogations afford states the opportunity to adjust to these changing conditions [@HBHelferFariss]. When states derogate from their international legal obligations, the derogation provides vital information to both international and domestic monitoring bodies, interest groups, and advocates about which rights are suspended, for how long, and the reasoning behind these suspensions. This information allows these actors—at least in principle—to challenge measures that are excessive, vague, or outlast the intended time frame of their implementation [@helfer2021].

Many legal critical assessments of derogations find importance in their use [@richardson2020], with some arguing that, "When a state does not bother with derogation provisions in times of emergencies where human rights are suspended, then it is a situation of unbridled state power” [@burchill2005, 97]. An abuse of state power can also potentially lead to the violation of non-derogable rights or rights whose protections cannot be affected by the declaration of a de jure emergency [@Greene:2020, 36]. These include rights such as the right to life, freedom from torture, freedom from slavery, freedom of thought and conscience, among others, that are viewed as essential rights that cannot be suspended even during emergencies. Research supports this notion—there is no meaningful variation in states' respect for derogable or non-derogable rights during officially declared states of emergencies [@richardsclay2012; @neumayer2013]. Even states that officially derogate from their international legal obligations are more likely to violate physical integrity rights, which are non-derogable rights [@neumayer2013; @comstock2019].

These findings suggest that states may not be discerning the significance of derogable vs. non-derogable rights status during emergencies. This also raises the question of the extent to which this distinction is valuable during the pandemic. Most derogations to the ICCPR have been for internal conflict or natural disaster. Is a public health emergency different from other states of emergency? Overall, there is a general expectation that following derogations, we may observe some decline in respect for human rights. However, previous studies have not explored the relationship between derogations and public health emergencies.

## Derogations during COVID-19

A first question of interest is the nature of states that are more likely to submit them in the first place. Little research has focused on derogations in political science, and research that does exist has mostly focused on the determinants of submitting derogations to the UN rather than the effects of these derogations. This literature tells us that states that are more democratic are more likely to issue derogations and other post-commitment actions such as reservations, understandings, and declarations [@Simmons2009; @comstock2019] and that most states facing an emergency do not, in fact, issue derogations [@HBHelferFariss]. The expectation is that states with strong democratic institutions need to convince domestic audiences of the legality and legitimacy of suspending rights during emergencies. Additionally, democratic states understand international law and international legal mechanisms to be more legitimate and recognize the importance of them more so than other states. Less democratic states, even though at times may be more likely to violate or evade international human rights law, are less likely to see nuanced pieces of treaty law as important to engage with. Scholars were split on expectations about COVID-era derogations. @Lebret:2020 argued that derogations are the best legal mechanism for states to notify the global community of their inability to comply with human rights standards during the COVID-19 pandemic. However, @richardson2020 see derogations as a less clear path for states during COVID-19, describing the actions, their conditions, and human rights as a "tangled morass" (p. 106).

```{r fig-derogation-map, fig.width=6.5, fig.height=3.5}
#| include: true
#| echo: false
#| fig-cap: "Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures"
ggplot() +
  geom_sf(data = derogation_map_data, aes(fill = derogations_1plus),
          linewidth = 0.1, color = "white") +
  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
  scale_fill_binned(low = colorspace::lighten(clrs[1], amount = 0.8), 
                    high = clrs[1], na.value = "grey85",
                    breaks = c(1, 5, 10, 15, 20),
                    limits = c(1, 20),
                    name = "Derogations filed: ",
                    guide = guide_colorsteps(barwidth = 7, barheight = 0.4,
                                             title.vjust = 1)) +
  theme_pandem() +
  theme(panel.border = element_blank(),
        legend.title = element_text(face = "bold"),
        legend.text = element_text(size = rel(0.6)),
        legend.justification = "center")
```

```{r tbl-derogation-countries}
#| tbl-cap: Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures 
#| include: true
#| echo: false
bind_cols(dplyr::slice(map_data_nice, 1:12),
          add_row(dplyr::slice(map_data_nice, 13:n()))) %>% 
  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]",
      col.names = rep(c("Country", "Derogations"), 2)) %>% 
  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))
```

We find that between January 2020 and June 2021, `r derog_detail$countries` states issued a total of `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries). Of derogations that detailed specific obligation suspension lengths, the average length was 41 days. Descriptively, we see states across multiple regions derogate, with notable concentration in South America, which has a robust history of engaging with human rights law. We also note that it is not necessarily the "usual suspects" of strong democracies that derogate. Notably, Nordic states—which typically are very active in human rights law engagement—did not issue any derogations during this period.

## Derogations from the ICCPR

In this research note, we look at states' derogations from the International Covenant on Civil and Political Rights (ICCPR) during the pandemic. The ICCPR is a foundational human rights treaty with broad human rights coverage. Exceptional circumstances brought on by the global pandemic have led to numerous, extensive derogations to the ICCPR removing state obligations to important rights areas of assembly, movement, and others. The ICCPR was adopted by the UN General Assembly on December 19, 1966, and entered into force on March 23, 1976. It was designed to recognize "the inherent dignity and of the equal and inalienable rights of all members of the human family" and outlined dozens of articles aimed at creating conditions where everyone can enjoy their civil, political, economic, social, and cultural rights. The ICCPR is one of only a few select human rights treaties that require states to report on the suspension or derogation of certain civil and political rights only under specific situations of "public emergency which threatens the life of the nation" [@iccpr]. The other treaties with this requirement include the European Convention Protection of Human Rights and Fundamental Freedoms and the American Convention on Human Rights. However, the ICCPR is the only one with global, rather than regional, membership and participation to track during the pandemic. As of 2023, there are 173 states that are party to the ICCPR. This means that most, but not all the UN’s 193 members had the ability to submit derogations during COVID-19. Other core human rights treaties do not contain the same requirement to submit derogation notifications whenever an anticipated violation will occur.

To legally derogate, governments must first officially declare a state of emergency (Article 4). Derogations must fulfill three criteria: they should be temporary in application, proportional to a crisis' actual threat and non-discriminatory. However, it is worth noting that states cannot derogate from the entirety of the ICCPR. Seven rights enshrined in the ICCPR are non-derogable: the right to life, the prohibition of torture or cruel, inhumane punishment, the prohibition of slavery, the right to recognition as a person before the law, and the protection of freedom of thought, conscience and religion. The international legal community viewed these rights as so fundamental and important that no type of crisis could justify their suspension. However, other rights can be suspended during crises. Several ICCPR articles provide that public health needs can justify limitations on or derogations from certain rights [@richardson2020, 112]: freedom of movement (Article 12), freedom of expression (Article 19), and the rights to peaceful assembly and freedom of association (Article 22).

Recognizing that public health exigencies would require states to curb civil and political rights, the UN Human Rights Office of the High Commissioner (OHCHR) published guidance stating that any emergency measures must meet specific criteria of legality, necessity, proportionality, and non-discrimination. However, not every country formally derogated—a number of countries implemented emergency measures without formally derogating from the ICCPR. Over 100 countries issued emergency declarations while fewer than 30 submitted COVID-19 related derogations [@icnl-covid]. These COVID-19 emergency declarations may have longer term impacts beyond temporary human rights violations [@BrownBrechenmacherCarothers:2020]. Below, we first theorize how derogations led to variation in the implementation of emergency measures during the pandemic, followed by their impact on human rights.

# The impact of derogations during COVID-19

## Derogations and emergency policies

During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21). As our descriptive data shows, not all states issued derogations. Derogating states were in the minority of states during the pandemic.

We expect that states that derogate during COVID-19 with the intention to impose emergency policies will follow through. Our expectation comes from the specific, time-limited nature of derogations and the limited audience they speak to. Critics of international law argue that states ratify treaties with the expectation of reputational and other benefits. However, derogations have no such public audience and are not signaling an increased commitment to human rights. There is no clear or expected connection with foreign aid, trade, or other theorized 'rewards' for larger legal actions. Instead, derogations are a very specific legal action that states voluntarily enter into after commitment has fully taken place.  Given this, we expect that derogating states took the legal action seriously and generally followed through with the derogations they submitted.

> **Hypothesis 1~a~**: If states issued COVID-19 derogations, then we expect that they will be more likely to impose relevant emergency polices such as canceling public events, restricting gatherings, closing public transit, and restricting domestic and international movement.
>
> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose relevant emergency policies.

## Derogations and human rights

If states determine that emergency measures are justified without regard for international law, the processes for respecting rights during crises and restoring rights at the end of the emergency may not be followed. As such, many states used the pandemic as an excuse to centralize authority and implemented measures that did not meet international legal standards. Many government engaged in discriminatory and abusive policy enforcement—they imposed restrictions on travel or movement by minority or opposition groups. For instance, in Australia, indigenous and migrant communities were disproportionately targeted in the enforcement of movement restrictions [@ICRC2021]. In Bulgaria, checkpoint controls went into effect in neighborhoods largely composed of the Roma community [@bulgaria2020]. In India, the police used a ban on public gatherings as an excuse to get rid of a months-long sit-in protesting the country's new citizenship law seen as discriminatory towards Muslims [@bbc2020].

The pandemic also led to the violation of non-derogable rights in many states. For instance, Greek officials intercepted and turned back boats filled with asylum-seekers and summarily returned asylum-seekers at the land border with Turkey. In doing so, they violated the non-derogable principle of non-refoulement—Article 7 of ICCPR prohibits people being sent back to another state where they are in danger of ill-treatment or torture [@HathawayStevensLim:2020]. Many other states changed rules of confinement affecting detainees, which also potentially violated non-derogable rights. For instance, in March 2020, a French government order extended the duration of pre-trial detention, raising the issue of prison overcrowding and leading to inhumane and degrading treatment [@Lebret:2020]. As such, this was not just a violation of a non-derogable right, but also raised questions regarding the proportionality of emergency measures.

We expect that states that did not formally derogate and did not notify the ICCPR of impending rights suspensions were less likely to take the criteria for derogations—necessity, proportionality, non-discrimination and time delimited—seriously, and were therefore more likely to impose measures that violated these criteria. If states determined that emergency measures are justified without consideration of international law, we expect that there will be more violations.

> **Hypothesis 2~a~**: If states formally derogate from international treaty provisions, then they are less likely to use discriminatory policy measures during the pandemic.
>
> **Hypothesis 2~b~**: If states formally derogate from international treaty provisions, then they are less likely to violate non-derogable rights.
>
> **Hypothesis 2~c~**: If states formally derogate from international treaty provisions, then they are less likely to impose measures with no time limits.

# Data and methods

```{r extract-details}
info <- list(
  n_weeks = length(unique(weekly_panel$year_week)),
  n_countries = length(unique(weekly_panel$iso3)),
  n_rows = label_comma()(nrow(weekly_panel)),
  dates = map(list(
    start_date = min(weekly_panel$year_week_day),
    end_date = max(weekly_panel$year_week_day)
  ), ~list(date = ., 
           nice = format(., "%B %e, %Y"),
           year = year(.)))
)
```

While some studies on repression, human rights, and the pandemic have focused within single country cases or across a region, we use global data to examine the relationship more broadly between treaty derogations, human rights, and state actions during the pandemic. We explore our hypotheses using data on COVID-19 derogations, state actions, health trends, and civil society measures that we collected from several different sources (see @tbl-data-sources), including start and end dates of ICCPR derogations [@UnitedNations:2023], formal government responses from the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021], indicators of the severity of government responses from the Varieties of Democracy project [@vdem-v12; @vdem-pandem], and COVID case and death counts from the World Health Organization's COVID-19 dashboard [@WHO:2023]. We include data from `r info$n_countries` countries during the first 15 months (or `r info$n_weeks` weeks) of the pandemic, from March 11, 2020 to June 30, 2021. While the Oxford Government Response Tracker provides daily data on a variety of government actions, submitting formal treaty responses to the ICCPR is less likely to be as immediate of an action as implementing domestic emergency actions like stay-at-home orders. Accordingly, we collapse this daily data into weekly totals, resulting in a panel with `r info$n_rows` country-week observations.

```{r tbl-data-sources}
#| tbl-cap: Sources for outcomes and explanatory variables
#| include: true
#| echo: false
tribble(
  ~`Outcome variables`,                        ~Description,                                                        ~Frequency,  ~Source,
  "Emergency public health measures",          "Binary indicator",                                                  "Weekly",    "Oxford COVID-19 Government Response Tracker",
  "Human rights and policy outcomes",          "Ordered categories for the severity of abuses",                     "Quarterly", "Varieties of Democracy PanDem",
  "ICCPR derogation",                          "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
  "New and cumulative COVID deaths and cases", "Counts",                                                            "Weekly",    "World Health Organization",
  "Past ICCPR derogation or action",           "Binary indicator",                                                  "Weekly",    "UN Treaty Collection",
  "Rule of law index",                         "0–1; higher values represent greater respect for rule of law",      "Annual",    "Varieties of Democracy",
  "Civil liberties index",                     "0–1; higher values represent greater respect for civil liberties",  "Annual",    "Varieties of Democracy",
  "Core civil society index",                  "0–1; higher values represent better environment for civil society", "Annual",    "Varieties of Democracy"
) %>% 
  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]") %>% 
  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA)) %>% 
  pack_rows(index = c(" " = 2, "Primary explanatory variable" = 1, "Secondary explanatory variables" = 5),
            indent = FALSE) %>% 
  column_spec(c(1, 2, 4), width = "0.25\\\\textwidth") %>% 
  column_spec(3, width = "0.1\\\\textwidth")
```

## Outcome variables

We examine the relationship between formal treaty derogations and different policy and human rights outcomes in two stages (see @tbl-vars). First, to test our first set of hypotheses and explore whether derogations were associated with emergency state responses in general, we look at the connection between ICCPR derogations and five different emergency public health measures recorded by the Oxford COVID-19 Government Response Tracker [@HaleAngristGoldszmidt:2021]: (1) cancelling public events, (2) imposing restrictions on gathering, (3) closing public transit, (4) restricting internal movement, and (5) limiting international travel. Each of these outcomes is recorded as an ordered categorical variable, with levels ranging from "no measures" to increasing counts and severity of emergency measures.[^severity] As we are most interested in the overall association between derogations and emergency measures, we collapse these variables into binary outcomes that indicate whether a country had no emergency measures or some emergency measures in each week.

[^severity]: Importantly, the Oxford Government Response Tracker measures the *stringency* or strictness of policy enforcement, not the *quality* of policies, so the values do not necessarily reflect how well countries implemented these emergency measures.

\begingroup\footnotesize

| General question                           | Outcomes                                                                                                                                  | Main explanatory variable  | Other explanatory variables                                                                                                                                              |
|-----------------|-----------------|-----------------|---------------------|
| Derogations and emergency measures (H~1~)  | Cancelling public events, Gathering restrictions, Close public transit, Restrictions on internal movement, Limits on international travel | ICCPR derogation in effect | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions                                                                     |
| Derogations and human rights (H~2~)        | Discriminatory policy, Violations of non-derogable rights, No time limit measures, Abusive enforcement                                    | ICCPR derogation in effect | New and cumulative COVID cases, new and cumulative COVID deaths, prior ICCPR derogations and actions, rule of law index, civil liberties index, core civil society index |

: Overview of outcomes and explanatory variables included in models {#tbl-vars}

\endgroup

Next, to test our second set of hypotheses, we explore the relationship between formal treaty derogations on four different human rights and policy outcomes from V-Dem's PanDem project [@vdem-pandem]: (1) whether pandemic-related emergency measures violated the ICCPR's principle of non-discrimination, (2) whether countries derogated from non-derogable ICCPR rights, (3) whether emergency measures failed to state a time limit, and (4) whether state security forces violated physical integrity rights while enforcing emergency measures. Similar to the Oxford data, each of these outcomes is an ordered category measuring the severity of the outcomes, including no, minor, moderate, and major violations. We do not collapse these variables into binary indicators and instead work with them as ordered categories. Importantly, since these PanDem variables are all measured quarterly, while our main data is recorded weekly, these variables act more like country-level fixed effects in our models, capturing the overall level of the severity of pandemic response each week.

\newpage

## Explanatory variables

### Key explanatory variable

For both sets of hypotheses, we explain changes in policy and human rights outcomes using a binary indicator of whether a state was under ICCPR derogation conditions during a given week. We collected and coded this data from real time state action updates recorded at the UN Treaty Collection [@UnitedNations:2023]. During the 15 months covered in this project, there were 115 formal derogations submitted mentioning 8 specific ICCPR articles; 15 derogations made no reference to any specific article. Of the derogations that indicated specific time limits, the average derogation duration was 41 days. We mark the start and end date of each derogation so that each country-week observation indicates whether a state has derogated.

### Secondary explanatory variables

We include several additional variables that are associated with and potentially confound both the decision to derogate and the decision to enact specific policies or human rights-related actions. To account for the weekly and overall severity of the pandemic in each country, we control for the count of new COVID-19 cases and the count of new COVID-19 deaths, as well as the cumulative count of cases and deaths.

Since formally derogating from treaties can serve as a signal that countries take their obligations seriously, we include an indicator marking whether a country has ever derogated or taken other formal ICCPR post-commitment actions, such as RUDs, prior to the pandemic. Some states new to derogations introduced them during the COVID-19 pandemic [@Comstock:2023]. However, past treaty action behavior may help explain entering derogations during this time. To examine this first step, we code two binary variables: (1) past ICCPR derogation behavior before 2020 and (2) past ICCPR post-commitment action behavior before 2020. This allows us to empirically test for the first step of engagement before analyzing any impact that COVID-19-era derogations had on different rights practices. There were 36 states that submitted ICCPR derogations and 60 that submitted any type of post-commitment action between 1966-2019. Finally, we include several annual measures of human rights, democracy, and civil society: V-Dem's rule of law index, civil liberties index, and core civil society index.

## Modeling strategy

Our data is structured as a balanced time series cross-sectional (TSCS) panel that has a natural hierarchical structure, with `r info$n_weeks` weekly observations nested within `r info$n_countries` countries. Accordingly, we model our outcome variables with multilevel or hierarchical models, with region-specific random offsets for each intercept and a fixed global year trend. When working with binary outcomes like the declaration of general emergency measures or the violation of non-derogable rights, we use logistic regression models; when working ordered categorical outcomes like the degree of discriminatory policy or abusive enforcement, we use ordered logistic regression models. We take a Bayesian approach to better model and explore the uncertainty inherent in measuring the effect of derogations on these broad policies.

We include complete formal specifications of all our models, likelihoods, and priors in the appendix. Here we include two simplified model specifications to help demonstrate the intuition behind these multilevel TSCS models. For binary outcomes, we use a logistic model with country-specific offsets:

::: {.column-page-inset-right}

$$
\begin{aligned}
&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + & \text{Probability of outcome} \\
&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random region offsets} \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-region variability,}
\end{aligned}
$$

:::

\noindent while for ordered outcomes, we use ordered logistic regression with country-specific offsets:

::: {.column-page-inset-right}

$$
\begin{aligned}
&\ \mathrlap{\textbf{Outcome level $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + & \text{Cumulative probability of outcome} \\
&\ \beta_{2 \dots n-1}\ \text{Other controls}_{it}\ + \beta_{n}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) & \text{Random region offsets} \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots n} \sim&\ \operatorname{Student\,t}(\nu = 1, \mu = 0, \sigma = 3) & \text{Prior for global averages and effects} \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 & \text{Prior for between-region variability} \\
\alpha_k \sim&\ \mathcal{N}(0, 1) & \text{Boundaries between thresholds}
\end{aligned}
$$

:::

In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society.

Because these models provide coefficients on the logged odds scale, and because the ordered logistic models require the incorporation of coefficients for each threshold to correctly interpret, we do not discuss the results of the models using raw model estimates (though we include complete regression tables in the appendix). Instead, we rely on probability-scale predictions and marginal effects whenever possible. To do so, we calculate conditional predicted probabilities holding all explanatory variables at their typical values (median or mode), and setting all region-specific random offsets ($b_{0_j}$) to 0, meaning that the predicted values refer to the effect of derogations in a typical region. For logistic models, we estimate the probability of having a specific policy or violating a specific human right; for ordered models, we estimate the probability of each possible level of severity. To determine the effect of derogations, we calculate the contrast in predicted probabilities when derogation is set to true vs. when it is set to false (e.g., in ordered logistic models we compare the probability of major discrimination under derogation with the probability of major discrimination without derogation). We use median values from our models' posterior distributions as point estimates and provide credible intervals using the 95% highest posterior density. For hypothesis testing, we report the posterior probability that the between-level predicted differences are above or below zero.

# Analysis

## Derogations and emergency policies

We begin by exploring our first set of hypotheses to examine how derogating and non-derogating states approached the implementation of their emergency public health measures. @fig-preds-mfx-policies-1 shows the results of five logistic regression models as predicted probabilities of imposing specific emergency measures over the course of the first 15 months of the pandemic. In every case, throughout the early pandemic, countries with a formal derogation in effect are far more likely to have undertook each policy, and the difference in probabilities between derogating and non-derogating states is substantial (see @fig-preds-mfx-policies-2 for the percentage-point differences between the predictions for the two groups of states).

For instance, all countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance. Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021. Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability. The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice` to `r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`). The differences in policy implementation are starker for other types of emergency measures. Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between `r policies_pred$gathering_restrictions.No.min$draw_nice`% and `r policies_pred$gathering_restrictions.No.max$draw_nice`%, representing a `r policies_mfx$gathering_restrictions.min$draw_nice`–`r policies_mfx$gathering_restrictions.max$draw_nice` percentage point difference (`r policies_mfx$gathering_restrictions.max$p_gt`). Derogating states were also far more likely to close public transit, beginning with a `r policies_pred$close_public_transit.Yes.max$draw_nice`% probability in March 2020 and ending in June 2021 with a `r policies_pred$close_public_transit.Yes.min$draw_nice`% probability. Non-derogating states followed a similar downward trend over time, but `r policies_mfx$close_public_transit.min$draw_nice`–`r policies_mfx$close_public_transit.max$draw_nice` percentage points lower than their derogating counterparts (`r policies_mfx$close_public_transit.max$p_gt`). Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to `r policies_pred$movement.Yes.min$draw_nice`% by June 2021. Non-derogating states, on the other hand, began the pandemic with a predicted `r policies_pred$movement.No.max$draw_nice`% probability of limiting movement, dropping to only `r policies_pred$movement.No.min$draw_nice`% by June 2021, representing a `r policies_mfx$movement.min$draw_nice`–`r policies_mfx$movement.max$draw_nice` percentage point difference (`r policies_mfx$movement.max$p_gt`). The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel. Both derogating and non-derogating states had a nearly 100% probability of limiting international movement across all 15 months.

```{r fig-preds-mfx-policies, fig.width=12, fig.height=6, out.width="100%", cache=FALSE}
#| include: true
#| echo: false
#| layout-ncol: 1
#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
#| fig-subcap:
#|   - "Predicted probabilities"
#|   - "Contrasts between predicted probabilities"

policies_plots <- policies_plot_data %>% 
  mutate(pred_plot = map2(pred_data, nice, ~{
    .x %>% 
      ggplot(aes(x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
                 color = derogation_ineffect, fill = after_scale(color), 
                 group = forcats::fct_rev(ordered(.width)))) +
      geom_lineribbon(alpha = 0.3) +
      scale_color_manual(values = c(clrs[1], clrs[8])) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      scale_y_continuous(labels = label_percent()) +
      labs(x = NULL, y = "Predicted probability", 
           color = "Derogation in effect", fill = "Derogation in effect",
           title = .y) +
      theme_pandem()
  })) %>% 
  mutate(mfx_plot = map2(mfx_data, nice, ~{
    .x %>% 
      mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>% 
      ggplot(aes(x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
                 group = forcats::fct_rev(ordered(.width)))) +
      geom_lineribbon(alpha = 0.25, fill = clrs[5], color = clrs[5]) +
      scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
      labs(x = NULL,
           y = "Percentage point difference\nin probability of policy",
           title = .y) +
      theme_pandem()
  }))

wrap_plots(policies_plots$pred_plot)
wrap_plots(policies_plots$mfx_plot)
```

These results reveal important insights into the motivations of derogating states. Countries that undertook emergency safety measures seem to have taken the time to follow international law and uphold their treaty obligations. International law thus appears to not just be an empty commitment—states that took the time to formally derogate were more likely to keep these emergency measures in place for longer and take their derogations seriously.

## Derogations and human rights

The results of our models show that derogations have some protective effect against more bureaucratic forms of human rights abuses and less against practical abuses (see @fig-preds-mfx-human-rights). There is no substantial difference between derogating and non-derogating countries when considering the probability of imposing discriminatory policies or engaging in abusive enforcement. Both types of countries follow nearly identical trends over the first 15 months of the pandemic, with an 85% predicted probability of no discriminatory policy in March 2020 rising to nearly 100% by June 2021. The next most probable outcomes—minor and moderate levels of discriminatory policies—are equally likely regardless of derogation status. Moreover, countries that derogate were substantially less likely to engage in no abusive enforcement and instead turn to minor or moderate levels of abusive enforcement, particularly in the early months of the pandemic. Minor and moderate levels of excessive enforcement are each several percentage points more likely for derogating countries than for their non-derogating counterparts.

```{r fig-preds-mfx-human-rights, fig.width=13, fig.height=6, out.width="100%"}
#| include: true
#| echo: false
#| layout-ncol: 1
#| fig-cap: "Predicted effects of imposing specific emergency public health measures over first 15 months of the COVID pandemic, split by whether states formally derogated from the ICCPR"
#| fig-subcap:
#|   - "Predicted probabilities"
#|   - "Contrasts between predicted probabilities"

human_rights_plots <- human_rights_plot_data %>% 
  # Predicted probabilities
  mutate(pred_plot = pmap(list(pred_data, nice, family), \(.data, .nice_name, .family) {
    if (.family == "cumulative") {
      .data %>% 
        ggplot(aes(
          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
          color = .category, fill = after_scale(color),
          group = forcats::fct_rev(ordered(.width))
        )) +
        geom_lineribbon(alpha = 0.3) +
        scale_color_manual(values = c(clrs[c(2, 4, 6, 7)])) +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        scale_y_continuous(labels = label_percent()) +
        labs(x = NULL, y = "Predicted probability", 
             color = "Severity", fill = "Severity",
             title = .nice_name) +
        facet_wrap(vars(derogation_ineffect)) +
        theme_pandem()
    } else {
      .data %>% 
        ggplot(aes(
          x = year_week_day, y = .epred, ymin = .lower, ymax = .upper,
          color = derogation_ineffect, fill = after_scale(color),
          group = forcats::fct_rev(ordered(.width))
        )) +
        geom_lineribbon(alpha = 0.3) +
        scale_color_manual(values = c(clrs[1], clrs[8])) +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        scale_y_continuous(labels = label_percent()) +
        labs(x = NULL, y = "Predicted probability", 
             color = "Derogation in effect", fill = "Derogation in effect",
             title = .nice_name) +
        theme_pandem()
    }
  })) %>% 
  # Marginal effects / contrasts
  mutate(mfx_plot = pmap(list(mfx_data, nice, family), \(.data, .nice_name, .family) {
    if (.family == "cumulative") {
      .data %>% 
        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
        ggplot(aes(
          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
          color = group, fill = after_scale(color),
          group = forcats::fct_rev(ordered(.width))
        )) +
        geom_lineribbon(alpha = 0.3) +
        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
        scale_color_manual(values = darken(clrs[c(2, 4, 6, 7)], 0.5, space = "combined"), guide = "none") +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        labs(x = NULL,
             y = "Percentage point difference\nin probability of outcome level",
             title = .nice_name) +
        facet_wrap(vars(group), ncol = 2) +
        theme_pandem()
    } else {
      .data %>% 
        mutate(across(c(draw, .lower, .upper), ~ . * 100)) %>%
        ggplot(aes(
          x = year_week_day, y = draw, ymin = .lower, ymax = .upper,
          group = forcats::fct_rev(ordered(.width))
        )) +
        geom_lineribbon(alpha = 0.3, fill = clrs[5], color = clrs[5]) +
        geom_hline(yintercept = 0, linewidth = 0.5, color = "grey50") +
        scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
        labs(x = NULL,
             y = "Percentage point difference\nin probability of outcome",
             title = .nice_name) +
        theme_pandem()
    }
  }))

wrap_plots(human_rights_plots$pred_plot)
wrap_plots(human_rights_plots$mfx_plot)
```

The mechanisms behind this seeming reversal—where countries that follow correct human rights procedures end up enforcing emergency policies in a more abusive way may be attributable to principal-agent problems. While elites or diplomats are the ones submitting derogations and committing to complying with international law, they are not the actors implementing emergency measures. During the pandemic, there may be immense variation in states' abilities to enforce these policies as intended, especially as it relates to human rights. Lower-level officials such as those in the police, state security forces and regional bureaucrats are often the ones enforcing emergency measures. Unlike their principals (elites), these agents neither have training in international law nor incentives to comply with derogations. Further, as previous research shows, during high levels of stress and fear, decision making is more likely to be driven by agents who discounts principals' (dis)incentives [@HooverGreen:2016]. As such, these differences in incentives and training between elites and local officials may explain the incidences of abusive enforcement even when states submit derogations and claim that emergency measures will be implemented proportional to the threat at hand.

In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status. For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability). The middle ground of only minor issues with time limits is exceptionally unlikely. Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues. The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish. Additionally, the risk of violating non-derogable rights is generally slightly lower in non-derogating states—countries that derogate are a percentage point less likely to violate non-derogable rights ($p(\Delta < 0) \approx 0.83$)

Derogations therefore appear to help protect against limitless emergency measures and somewhat against violations of non-derogable rights, but they do not appear to do much to protect against actual observed human rights abuses like discriminatory policies and abuse enforcement of those policies. We thus find strong support only for H~2c~—that derogating states are less likely to impose measures with no time limits. We find weak support for our hypothesis that derogating states will be less likely to violate non-derogable rights (H~2b~) and no support for our hypothesis that derogations protect against discriminatory policies (H~2c~). Derogations seem to help with bureaucratic processes and might protect against severe violations of human rights, but they do not appear to influence the overall implementation of emergency measures.

# Conclusion

The COVID-19 pandemic raised numerous concerns that the crisis would contribute to violations of democratic standards and human rights in a number of countries and would reinforce the powers of repressive governments worldwide. Therefore, this research note set out to examine state responses to their international human rights treaty obligations, particularly the ICCPR, during the COVID-19 pandemic and any subsequent impact on human rights violations. We focused specifically on the impact of derogations—an international legal means to communicate to both the UN and the global community that certain rights will be suspended during times of crisis. Derogations are meant to provide states with flexibility during crises and to encourage them to sign and ratify treaties without concern about the inability to adapt to times of emergency. International legal scholarship has been divided over expectations of derogations during the pandemic.

With the caveat that our results are not causal—we have not attempted to use any formal strategy to identify an unconfounded causal effect, and our findings are therefore descriptive—we find that there is good news for compliance and derogations. Derogators did not systematically use this legal "escape clause" opportunistically—derogators used the process as intended. Most derogations included time limitations, states tended to adhere to the time limits, and there was not an overall abuse of the derogation time to abuse, repress, and discriminate. Though most states did not submit COVID-19 related derogations, the ones that did used derogations as intended by international law. Once the derogation ended, states readjusted to typical compliance once more. States therefore complied with the intent of derogations and the purpose of the treaty. This is an important finding for international legal behavior and compliance studies. Even in a time of crisis, overall, states considered it important to follow through with the processes and expectations of international human rights law bureaucracy.

We suggest that future research examine different incentives facing state officials during crisis. There may be key differences in enforcement between local level bureaucrats and higher-level elites. While elites submit derogations and may be more cautious about international law, local level enforcement may not be as concerned about international legal aspects such as time limits and derogable vs. non-derogable rights. This distinction may explain some variation in policy and discrimination practices during the pandemic. We also recommend that future research analyze the impact of other factors that can mitigate the adverse impacts of emergency measures. In the absence of adequate review from the international community and treaty monitoring bodies during crises, certain domestic-level factors, such as a robust civil society and media, may be necessary to mitigate the adverse effects of emergency measures.

Overall, we find that states used derogations as they were intended during the COVID-19 pandemic and no large-scale misuse of the legal action took place. It can be an important signal for states to submit derogations, both to international and domestic communities, about intentions to mitigate the spread of disease through the temporary suspension of human rights.

# References