### 2023-05-05 manuscript/appendix.qmd
Add appendix

---
title: 'Online appendix for "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"'
short-title: 'Online appendix for "Pandemic Pass?"'
subtitle: "(Research note)"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  orcid: 0000-0003-3506-7590
  url: https://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Audrey L. Comstock
  affiliation: Arizona State University
  orcid: 0000-0003-1772-5350
  url: https://audreylcomstock.weebly.com/
  email: audrey.comstock@asu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  orcid: 0000-0002-3948-3914
  url: https://www.andrewheiss.com
  email: aheiss@gsu.edu
date: May 2, 2023
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
bibliography: bib.bib
link-citations: true
crossref: 
  tbl-title: Appendix Table
  fig-title: Appendix Figure
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
                      warning = FALSE, message = FALSE)

knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE)

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r libraries-data}
library(tidyverse)
library(modelsummary)
library(scales)
library(kableExtra)
library(gt)
library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Generated via random.org
set.seed(8511)

tar_load(c(policies_table_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))

# modelsummary functions
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))

add_breaks <- function(x, fmt) {
  if (fmt == "latex") {
    kableExtra::linebreak(x, align = "c")
  } else (
    str_replace(x, "\\n", "<br>")
  )
}
```

# Modeling approach

We use Stan `r cmdstanr::cmdstan_version()` [@stan] through R `r paste0(version$major, ".", version$minor)` [@r-project] and {brms} `r packageVersion("brms")` [@brms-jss] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/MTR6X>. Our general modeling approach can be summarized as follows:

## Model definitions

### Logistic regression

$$
\begin{aligned}
&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
\end{aligned}
$$

### Ordered logistic regression

$$
\begin{aligned}
&\ \mathrlap{\textbf{Model of outcome level $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
&\ \textbf{Models for distribution parameters} \\
\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
\alpha_k \sim&\ \mathcal{N}(0, 1)
\end{aligned}
$$

# H~1~: Derogations and emergency policies

## Predicted probabilities and group contrasts

```{r tbl-policies-summary}
#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries
#| include: true
#| echo: false
tbl_part1 <- policies_table_data$policies_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, time, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num")

tbl_part2 <- policies_table_data$policies_mfx_tbl %>% 
  mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
  mutate(num = glue::glue("{draw_nice}\n({min_nice}–{max_nice})")) %>% 
  select(nice, year_week_day, diff_type, num, p_gt_0) %>% 
  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_week_day")) %>% 
  mutate(derogation_ineffect = "No")

tbl_part1 %>% 
  left_join(tbl_part2, by = join_by(nice, derogation_ineffect)) %>% 
  mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
  select(derogation_ineffect, `March 2020`, `June 2021`,
         `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
         `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`) %>% 
  kbl(col.names = c("Derogation", "March 2020", "June 2021",
                    "Week", "Difference", "p > 0", "Week", "Difference", "p > 0"),
      align = "lcccccccc",
      booktabs = TRUE, linesep = "",
      escape = FALSE) %>% 
  kable_styling(font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
  column_spec(3, border_right = T) %>% 
  pack_rows(index = table(fct_inorder(tbl_part1$nice))) %>% 
  add_header_above(c(" " = 1, "Predicted probabilities" = 2, 
                     "Smallest difference" = 3, "Largest difference" = 3))
```

## Complete model results

```{r tbl-results-full-policies}
#| tbl-cap: Results from models showing relationship between derogations and emergency policies
#| include: true
#| echo: false
notes <- c(
  "Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
  "95% credible intervals (highest density posterior interval, or HDPI) in brackets.",
  "Total \\(R^2\\) considers the variance of both population and group effects;",
  "marginal \\(R^2\\) only takes population effects into account."
)

if (fmt_out == "latex") {
  notes <- notes %>% 
    str_replace_all(r"[\\]", r"[\\\\]") %>% 
    str_replace_all("%", r"(\\\\%)")
} else {
  notes <- paste(notes, collapse = " ")
}

if (fmt_out == "latex") { 
  modelsummary(models_tbl_policies,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "kableExtra",
               fmt = fmt_significant(2),
               escape = FALSE) %>% 
    kable_styling(htmltable_class = "table-sm light-border",
                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
    column_spec(1:6, width = "0.16\\\\textwidth")
} else {
  modelsummary(models_tbl_policies,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "gt",
               fmt = fmt_significant(2)) %>% 
    tab_footnote(footnote = notes) %>% 
    tab_style(style = cell_text(weight = "bold"),
              locations = cells_column_labels())
}
```

# H~2~: Derogations and human rights

## Predicted probabilities and group contrasts

## Complete model results

```{r tbl-results-full-human-rights}
#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~1~)
#| include: true
#| echo: false
if (fmt_out == "latex") {
  modelsummary(models_tbl_human_rights,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "kableExtra",
               fmt = fmt_significant(2),
               escape = FALSE) %>% 
    kable_styling(htmltable_class = "table-sm light-border",
                  font_size = ifelse(fmt_out == "latex", 8, NA)) %>% 
    footnote(general = notes, footnote_as_chunk = FALSE, escape = FALSE) %>% 
    column_spec(1:5, width = "0.18\\\\textwidth")
} else {
  modelsummary(models_tbl_human_rights,
               estimate = "{estimate}",
               statistic = "[{conf.low}, {conf.high}]",
               coef_map = coef_map,
               gof_map = gof_map,
               output = "gt",
               fmt = fmt_significant(2),
               escape = FALSE) %>% 
    tab_footnote(footnote = notes) %>% 
    tab_style(style = cell_text(weight = "bold"),
              locations = cells_column_labels())
}
```

# References