### 2023-05-05 analysis/appendix.qmd
Add appendix to website

---
title: Appendix
toc: false
page-layout: full
---

```{r include=FALSE}
# Add the manuscript target as a dependency to the overall website
targets::tar_load(c(manuscript_nice, manuscript_manuscripty))
```

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/appendix.html" target="appendix">
  <i class="fa-solid fa-file-lines"></i> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/appendix.html" target="_blank">
  <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/appendix.pdf" target="appendix">
  <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
  </a>

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/appendix.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>

  <a class="btn btn-warning btn-sm mb-1" href="/manuscript/output-ms/appendix.pdf" target="appendix">
  <i class="fa-solid fa-file-pdf"></i>&ensp;Manuscripty PDF
  </a>

  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output-ms/appendix.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;Manuscripty PDF <small>(download)</small>
  </a>
</p>

<iframe src="/manuscript/output/appendix.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="appendix"></iframe>
```