### 2023-05-10 index.qmd

---
title: "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  orcid: 0000-0003-3506-7590
  url: https://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Audrey L. Comstock
  affiliation: Arizona State University
  orcid: 0000-0003-1772-5350
  url: https://audreylcomstock.weebly.com/
  email: audrey.comstock@asu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  orcid: 0000-0002-3948-3914
  url: https://www.andrewheiss.com
  email: aheiss@gsu.edu
format:
  html:
    code-tools: false
    toc: false
sidebar: false
---

```{r echo=FALSE}
yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
```

### Abstract

`r abstract`