### 2023-06-03 presentations/2022-11-04_peace-science/index.qmd
Better handling of slide resources

---
title: "Peace Science Society"
date: 2022-11-04
toc: false
abstract: |
  Annual conference of the Peace Science Society; Denver, Colorado
abstract-title: "Location"
published-title: "Presented"
image: first-slide.png
page-layout: full
---

<a class="btn btn-success btn-lg" target="_blank" href="slides/2022-11-04_peace-science.html">{{< fa arrow-up-right-from-square >}} &ensp;View all slides in new window</a>

<div class="ratio ratio-16x9">
<iframe class="slide-deck" src="slides/2022-11-04_peace-science.html"></iframe>
</div>