### 2024-05-16 README.qmd
Use new manuscript target name

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)

manuscript_target <- targets::tar_read(manuscript)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
```

# `r title`

<!-- badges: start -->
<!-- badges: end -->

[Suparna Chaudhry](https://www.suparnachaudhry.com/) • Lewis and Clark College  
[Audrey Comstock](https://audreylcomstock.weebly.com/) • Arizona State University  
[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

---

## Abstract

`r abstract`

---

## How to download and replicate

You can either [download the compendium as a ZIP file](/archive/main.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

We use the [{renv} package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [{targets} package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis. ([See this for a short helpful walkthrough of {targets}.](https://books.ropensci.org/targets/walkthrough.html)).

To reproduce the findings and re-run the analysis, do the following:

1. Download and install these fonts (if you’re using Windows, make sure you right click on the font files and choose “Install for all users” when installing these fonts):
    - [Noto Sans](https://fonts.google.com/specimen/Noto+Sans)
    - [Linux Libertine O](https://www.cufonfonts.com/font/linux-libertine-o) (also [here](https://sourceforge.net/projects/linuxlibertine/))
    - [Libertinus Math](https://github.com/alerque/libertinus)
2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library
    - If you’re using Windows, [install RTools too](https://cran.r-project.org/bin/windows/Rtools/) and add it to your PATH so that you can install packages from source if needed
3. Open `mountainous-mackerel.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
4. In the terminal, run `quarto install tinytex` to ensure that you have a working LaTeX installation.
5. If it's not installed already, R *should* try to install the {renv} package when you open the RStudio Project for the first time. If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
6. Run `renv::restore()` in the R console to install all the required packages for this project.
7. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.

Running `targets::tar_make()` will create several helpful outputs:

1. All project data in `data/`
2. An analysis notebook website in `_site/`
3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/`

## `r emoji_glue(":snow_capped_mountain::fish:")`: Note on "mountainous mackerel" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.4.0

codename(seed = "covid and ngos", type = "ubuntu")
#> [1] "mountainous mackerel"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.