### 2024-07-24 manuscript/appendix.qmd
Only set LIBGS on macOS

---
title: 'Online appendix for "Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19"'
short-title: 'Online appendix for "Pandemic Pass?"'
subtitle: "(Research note)"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  orcid: 0000-0003-3506-7590
  url: https://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Audrey L. Comstock
  affiliation: Arizona State University
  orcid: 0000-0003-1772-5350
  url: https://audreylcomstock.weebly.com/
  email: audrey.comstock@asu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  orcid: 0000-0002-3948-3914
  url: https://www.andrewheiss.com
  email: aheiss@gsu.edu
date: July 22, 2024
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/mountainous-mackerel>"
bibliography: references.json
csl: csl/chicago-author-date.csl
link-citations: true
crossref: 
  tbl-title: "Table&nbsp;A"
  fig-title: "Figure&nbsp;A"
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(echo = FALSE, include = FALSE, 
                      warning = FALSE, message = FALSE)

knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE)

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE)
```

```{r tikz-stuff, include=FALSE}
if (Sys.info()["sysname"] == "Darwin") {
  # Necessary for using dvisvgm on macOS
  # See https://www.andrewheiss.com/blog/2021/08/27/tikz-knitr-html-svg-fun/
  Sys.setenv(LIBGS = "/usr/local/share/ghostscript/9.53.3/lib/libgs.dylib.9.53")
}

# Conditional tikz output types; use PDF if this is LaTeX, otherwise use SVG with outlined fonts
if (fmt_out == "latex") {
  knitr::opts_template$set(
    tikz_settings = list(
      fig.ext = "pdf", 
      fig.align = "center", 
      extra.preamble = c(
        "\\usepackage{libertine}",
        "\\usepackage{libertinust1math}"
      )
    )
  )
} else {
  knitr::opts_template$set(
    tikz_settings = list(
      fig.ext = "svg", 
      fig.align = "center", 
      extra.preamble = c(
        "\\usepackage{libertine}",
        "\\usepackage{libertinust1math}"
      ),
      engine.opts = list(dvisvgm.opts = "--no-fonts")
    )
  )
}
```

```{r libraries-data}
library(tidyverse)
library(patchwork)
library(modelsummary)
library(scales)
library(tinytable)
library(targets)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Generated via random.org
set.seed(8511)

who_regions_raw <- tar_read(iccpr_who_clean) %>% 
  distinct(country_name, iso3, who_region, who_region_long)

tar_load(year_quarter_lookup)
tar_load(c(policies_table_data, hr_table_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))

# Plotting and modelsummary functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(modelsummary_functions), .GlobalEnv))

add_breaks <- function(x, fmt) {
  sapply(x, \(text) {
    if (is.na(text)) {
      return(NA)
    }

    if (fmt == "latex") {
      text <- str_replace_all(text, c("\\[" = "{[}", "\\]" = "{]}"))
      if (str_detect(text, "\\n")) {
        str_c("{", str_replace(text, "\\n", r"{ \\\\ }"), "}")
      } else {
        text
      }
    } else {
      str_replace(text, "\\n", "<br>")
    }
  }) %>% unname()
}
```

::: {.content-visible when-meta="list-of-figures-normal"}
```{=latex}
\listoftables
\listoffigures
```
:::

::: {.content-visible when-meta="list-of-figures-single"}

```{=latex}
\begin{singlespace}
\listoftables
\listoffigures
\end{singlespace}
```

:::

\clearpage

# Modeling approach

We use Stan `r cmdstanr::cmdstan_version()` [@stan] through R `r paste0(version$major, ".", version$minor)` [@rproject-4-3-3] and {brms} `r packageVersion("brms")` [@Burkner:2017] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis compendium at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW>.

## Priors

We follow the suggestion of @GelmanJakulinPittau:2008 and use weakly informative priors for our logistic and ordered logistic regression models. For consistency with prior specification, and for computation efficiency, we mean-center all nonbinary variables so that parameter estimates represent changes from the mean. We use two general priors (see @fig-priors):

- For all $\beta$ terms, we use a Student t distribution with a mean of 0 and a standard deviation of 3. This keeps most parameter estimates around −5 to 5, with thicker tails that allow for some possibility of extreme values.
- For $\sigma$ terms related to the variance or standard deviation of parameter distributions, which must be positive, we use a half Cauchy distribution, centered at 0 with a $\gamma$ of 1

These priors give more weight to realistic areas of parameter values and downweight values in unrealistic spaces. For instance, since logit-scale coefficient values greater than 4 or 5 are highly unlikely, our Student t prior puts more weight on smaller values. Additionally, weakly informative priors allow reasonable and considerable uncertainty in possible parameter estimates.

```{r}
#| label: fig-priors
#| fig-cap: "Density plots of prior distributions for model parameters"
#| fig-width: 6.5
#| fig-height: 2.5
#| out-width: "100%"
#| include: true
#| echo: false

p1 <- ggplot() +
  stat_function(geom = "area", 
                fun = ~extraDistr::dlst(., df = 1, mu = 0, sigma = 3), 
                fill = clrs[2]) +
  xlim(c(-20, 20)) +
  labs(x = "βs") +
  facet_wrap(vars("β: Student t(ν = 1, µ = 0, σ = 3)")) +
  theme_pandem(prior = TRUE)

p2 <- ggplot() +
  stat_function(geom = "area", 
                fun = ~dcauchy(., 0, 1), 
                fill = clrs[4]) +
  xlim(c(0, 10)) +
  labs(x = "σ") +
  facet_wrap(vars("σ: Cauchy(x = 0, γ = 1), lower bound = 0")) +
  theme_pandem(prior = TRUE)

plot_priors <- p1 | p2
plot_priors
# ggsave("submissions/figs/plot_priors.svg", plot_priors, width = 6.5, height = 2.5)
```

## Model definitions

### H~1~: Logistic regression

{{< include tikz/h1.md >}}

$$
\begin{aligned}
&\ \mathrlap{\textbf{Binary outcome $i$ across week $t$ within each country $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Bernoulli}(\pi_{it_j}) \\[0.75em]
&\ \textbf{Distribution parameters} \\
\pi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Week number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0
\end{aligned}
$$

The actual R code for these models is included in the replication code at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW> This is a simplified representation of the {brms} [@Burkner:2017] model code:

```r
# H1: Logistic regression for binary outcomes
brm(
  bf(outcome ~ derogation_ineffect + 
      new_cases_z + cumulative_cases_z + 
      new_deaths_z + cumulative_deaths_z +
      prior_iccpr_derogations + prior_iccpr_other_action + 
      v2x_rule + v2x_civlib + v2xcs_ccsi + 
      year_week_num + (1 | country_name)),
  family = bernoulli(),
  prior = c(
    prior(student_t(1, 0, 3), class = Intercept),
    prior(student_t(1, 0, 3), class = b),
    prior(cauchy(0, 1), class = sd, lb = 0)),
  ...
)
```

### H~2~: Ordered logistic regression

{{< include tikz/h2.md >}}

$$
\begin{aligned}
&\ \mathrlap{\textbf{Model of outcome level $i$ across quarter $t$ within each region $j$}} \\
\text{Outcome}_{it_j} \sim&\ \operatorname{Ordered\ logit}(\phi_{it_j}, \alpha_k) \\[0.75em]
&\ \textbf{Models for distribution parameters} \\
\phi_{it_j} =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Derogation in effect}_{it} + \\
&\ \beta_2\ \text{New cases}_{it}\ + \beta_3\ \text{Cumulative cases}_{it}\ + \\
&\ \beta_4\ \text{New deaths}_{it}\ + \beta_5\ \text{Cumulative deaths}_{it}\ + \\
&\ \beta_6\ \text{Past ICCPR derogation}_{it}\ + \beta_7\ \text{Past ICCPR action}_{it}\ + \\
&\ \beta_8\ \text{Rule of law index}_{it}\ + \beta_9\ \text{Civil liberties index}_{it}\ + \\
&\ \beta_{10}\ \text{Core civil society index}_{it}\ + \beta_{11}\ \text{Quarter number}_{it} \\
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \\[0.75em]
&\ \textbf{Priors} \\
\beta_{0 \dots 11} \sim&\ \operatorname{Student\ t}(\nu = 1, \mu = 0, \sigma = 3) \\
\sigma_0 \sim&\ \operatorname{Cauchy}(x = 0, \gamma = 1) \text{, lower bound} = 0 \\
\alpha_k \sim&\ \mathcal{N}(0, 1)
\end{aligned}
$$

The actual R code for these models is included in the replication code at <https://doi.org/10.17605/OSF.IO/ANONYMIZED-FOR-NOW> This is a simplified representation of the {brms} [@Burkner:2017] model code:

```r
# H2: Ordinal logistic regression for ordered outcomes
brm(
  bf(outcome ~ derogation_ineffect + 
      new_cases_z + cumulative_cases_z + 
      new_deaths_z + cumulative_deaths_z +
      prior_iccpr_derogations + prior_iccpr_other_action + 
      v2x_rule + v2x_civlib + v2xcs_ccsi + 
      year_quarter_num + (1 | who_region)),
  family = cumulative(),
  prior = c(
    prior(student_t(1, 0, 3), class = Intercept),
    prior(student_t(1, 0, 3), class = b),
    prior(cauchy(0, 1), class = sd, lb = 0)),
  ...
)
```

::: {.content-visible when-format="html"}
# H~1~: Derogations and emergency policies

## Predicted probabilities and group contrasts
:::

\blandscape

```{r tbl-policies-summary}
#| tbl-cap: Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for emergency policy models
#| include: true
#| echo: false

tbl_part1 <- policies_table_data$policies_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_week_day == "2020-03-09", "March 2020", "June 2021")) %>% 
  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
  select(nice, time, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num")

tbl_part2 <- policies_table_data$policies_mfx_tbl %>% 
  mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
  select(nice, year_week_day, diff_type, num, p_gt_0) %>% 
  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_week_day")) %>% 
  mutate(derogation_ineffect = "No")

tbl_groups <- tbl_part1 %>%
  select(nice) %>% 
  mutate(row = 1:n()) %>%
  distinct(nice, .keep_all = TRUE) %>% 
  mutate(row_cumulative = row + row_number() - 1)

tbl_part1 %>%
  left_join(tbl_part2, by = join_by(nice, derogation_ineffect)) %>%
  mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
  select(
    derogation_ineffect, `March 2020`, `June 2021`,
    `year_week_day_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
    `year_week_day_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`
  ) %>%
  set_names(
    c(
      "Derogation", "March 2020", "June 2021",
      "Week", "$\\Delta$", "p > 0", "Week", "$\\Delta$", "p > 0"
    )
  ) %>%
  tt(width = c(0.7, 1, 1, 1, 0.8, 0.5, 1, 0.8, 0.5) |> {\(x) x / sum(x)}()) %>%
  format_tt() %>%
  style_tt(j = 1:9, align = "lcccccccc") %>%
  style_tt(i = 1:20, j = 3, line = "r", line_width = 0.1, line_color = "#000000") %>%
  group_tt(
    i = pull(tbl_groups, row, name = nice) %>% as.list(),
    j = list(
      "Predicted probabilities" = 2:3, 
      "Smallest difference" = 4:6, 
      "Largest difference" = 7:9),
    indent = 0) %>% 
  style_tt(
    i = tbl_groups$row_cumulative, 
    bold = TRUE, background = "#e6e6e6", align = "l"
  ) %>% 
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = "rowsep=1pt, cells={font=\\scriptsize}, row{2}={valign=b}"
  )
```

\elandscape

::: {.content-visible when-format="html"}
## Complete model results
:::

```{r tbl-results-full-policies}
#| tbl-cap: "Complete results from models showing relationship between derogations and emergency policies (H~1~)"
#| include: true
#| echo: false

notes <- paste(
  "Note: Estimates are median posterior log odds from ordered logistic and binary logistic regression models;",
  "95% credible intervals (highest density posterior interval, or HDPI) in brackets.",
  "Total \\(R^2\\) considers the variance of both population and group effects;",
  "marginal \\(R^2\\) only takes population effects into account."
)

if (fmt_out == "latex") {
  notes <- str_replace_all(notes, "%", r"(\\%)")
}

inner <- "
row{odd[3-27]}={abovesep=0pt, belowsep=6pt},
row{even[2-26]}={rowsep=0pt},
cells={font=\\scriptsize}, row{1}={valign=b}
"

css <- "
.coef-table-1 tbody tr:not(:nth-child(n+26)):nth-child(odd) td {
    padding-bottom: 0;
}
.coef-table-1 tbody tr:not(:nth-child(n+27)):nth-child(even) td {
    padding-top: 0;
}
"

modelsummary(models_tbl_policies,
  estimate = "{estimate}",
  statistic = "[{conf.low}, {conf.high}]",
  coef_map = coef_map,
  gof_map = gof_map,
  output = "tinytable",
  fmt = fmt_significant(2),
  notes = notes,
  width = c(0.2, rep(0.16, 5)),  
) %>% 
  style_tt(i = seq(1, 25, 2), j = 1, rowspan = 2, alignv = "t") %>% 
  style_tt(
    bootstrap_class = "table table-sm coef-table-1",
    bootstrap_css_rule = css,
    tabularray_inner = inner
  )
```

::: {.content-visible when-format="html"}
# H~2~: Derogations and human rights

## Predicted probabilities and group contrasts
:::

\blandscape

```{r table-hr-summary}
#| include: true
#| echo: false

tbl_part1 <- hr_table_data$hr_pred_tbl %>% 
  mutate(across(c(draw, min, max), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(time = ifelse(year_quarter_day == "2020-01-01", "2020-Q1", "2021-Q2")) %>% 
  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
  select(nice, time, group = category, derogation_ineffect, num) %>% 
  pivot_wider(names_from = "time", values_from = "num") %>% 
  mutate(group = recode(group, "NA" = "—"))

tbl_part2 <- hr_table_data$hr_mfx_tbl %>% 
  left_join(year_quarter_lookup, by = join_by(year_quarter_day)) %>% 
  mutate(across(c(draw, min, max, p_gt_0), list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(group = recode(group, "Logit" = "—")) %>% 
  mutate(diff_type = ifelse(rank == "min", "Smallest difference", "Largest difference")) %>% 
  mutate(num = glue::glue("{draw_nice}\n[{min_nice}–{max_nice}]")) %>% 
  select(nice, year_quarter, group, diff_type, num, p_gt_0) %>% 
  pivot_wider(names_from = "diff_type", values_from = c("num", "p_gt_0", "year_quarter")) %>% 
  mutate(derogation_ineffect = "No")

tbl_groups <- tbl_part1 %>%
  select(nice) %>% 
  mutate(row = 1:n()) %>%
  distinct(nice, .keep_all = TRUE) %>% 
  mutate(row_cumulative = row + row_number() - 1)

tbl_part1 %>%
  left_join(tbl_part2, by = join_by(nice, derogation_ineffect, group)) %>%
  mutate(across(where(is.character), ~add_breaks(., fmt = fmt_out))) %>% 
  select(
    derogation_ineffect, group, `2020-Q1`, `2021-Q2`,
    `year_quarter_Smallest difference`, `num_Smallest difference`, `p_gt_0_Smallest difference`,
    `year_quarter_Largest difference`, `num_Largest difference`, `p_gt_0_Largest difference`
  ) %>%
  set_names(
    c(
      "Derogation", "Level", "2020-Q2", "2021-Q2",
      "Week", "$\\Delta$", "p > 0", "Week", "$\\Delta$", "p > 0"
    )
  ) %>% 
  tt(
    width = c(0.7, 0.7, 1, 1, 1, 0.8, 0.5, 1, 0.8, 0.5) |> {\(x) x / sum(x)}(),
    caption = "Summary of predicted probabilities and minimum and maximum contrasts between derogating and non-derogating countries for human rights models \\label{tbl-hr-summary}") %>%
  format_tt() %>%
  style_tt(j = 1:10, align = "lccccccccc") %>%
  style_tt(i = 1:30, j = 4, line = "r", line_width = 0.1, line_color = "#000000") %>%
  group_tt(
    i = pull(tbl_groups, row, name = nice) %>% as.list(),
    j = list(
      "Predicted probabilities" = 3:4, 
      "Smallest difference" = 5:7, 
      "Largest difference" = 8:10),
    indent = 0) %>% 
  style_tt(
    i = tbl_groups$row_cumulative, 
    bold = TRUE, background = "#e6e6e6", align = "l"
  ) %>% 
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_outer = "theme=long",
    tabularray_inner = "rowsep=1pt, cells={font=\\scriptsize}, row{2}={valign=b}"
  ) %>% 
  theme_tt("multipage", rowhead = 2)
```

\elandscape

::: {.content-visible when-format="html"}
## Complete model results
:::

```{r tbl-results-full-human-rights}
#| tbl-cap: Complete results from models showing relationship between derogations and human rights (H~2~)
#| include: true
#| echo: false

inner <- "
row{odd[3-33]}={abovesep=0pt, belowsep=2pt},
row{even[2-34]}={abovesep=0pt, belowsep=0pt},
cells={font=\\scriptsize}, row{1}={valign=b}
"

css <- "
.coef-table-2 tbody tr:not(:nth-child(n+32)):nth-child(odd) td {
    padding-bottom: 0;
}
.coef-table-2 tbody tr:not(:nth-child(n+33)):nth-child(even) td {
    padding-top: 0;
}
"

modelsummary(models_tbl_human_rights,
  estimate = "{estimate}",
  statistic = "[{conf.low}, {conf.high}]",
  coef_map = coef_map,
  gof_map = gof_map,
  output = "tinytable",
  fmt = fmt_significant(2),
  notes = notes,
  width = rep(0.2, 5),  
) %>% 
  style_tt(i = seq(1, 25, 2), j = 1, rowspan = 2, alignv = "t") %>% 
  style_tt(
    bootstrap_class = "table table-sm coef-table-2",
    bootstrap_css_rule = css,
    tabularray_inner = inner
  )
```

\clearpage

```{r tbl-who-regions}
#| tbl-cap: WHO regions
#| include: true
#| echo: false
who_regions <- who_regions_raw %>%
  mutate(
    region_title = glue::glue("{who_region}: {who_region_long}")
  ) %>% 
  group_by(region_title) %>% 
  summarize(
    countries = str_flatten(country_name, collapse = ", ", last = ", and ")
  ) %>% 
  pivot_longer(cols = everything()) %>% 
  select(-name) %>% 
  as.data.frame() %>%  # tibbles don't like NULL column names
  set_names(NULL)

who_regions %>% 
  tt(width = 1) %>% 
  format_tt(j = 1, escape = TRUE) %>% 
  style_tt(i = seq(1, nrow(who_regions), by = 2), bold = TRUE, background = "#e6e6e6") %>% 
  style_tt(i = seq(2, nrow(who_regions), by = 2), indent = 0, align = "l") %>% 
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = "rowsep=1pt, cells={font=\\footnotesize}"
  )
```

\clearpage

```{r tbl-achr-echr-iccpr}
#| tbl-cap: States that derogated from regional human rights conventions
#| include: true
#| echo: false

achr_echr_iccpr <- tribble(
                 ~ACHR, ~Also,             ~ECHR, ~Also1,
           "Argentina", "Yes",         "Albania",   "No",
             "Bolivia",  "No",         "Armenia",  "Yes",
               "Chile", "Yes",      "Azerbaijan",  "Yes",
            "Colombia", "Yes",         "Estonia",  "Yes",
  "Dominican Republic", "Yes",         "Georgia",  "Yes",
             "Ecuador", "Yes",          "Latvia",  "Yes",
         "El Salvador", "Yes",         "Moldova",  "Yes",
           "Guatemala", "Yes", "North Macedonia",   "No",
            "Honduras",  "No",         "Romania",  "Yes",
             "Jamaica",  "No",      "San Marino",  "Yes",
              "Panama",  "No",          "Serbia",   "No",
            "Paraguay", "Yes",                NA,     NA,
                "Peru", "Yes",                NA,     NA,
            "Suriname",  "No",                NA,     NA,
           "Venezuela",  "No",                NA,     NA
  )

achr_echr_iccpr %>% 
  as.data.frame() %>% 
  set_names(
    c(
      "States that derogated from the ACHR",
      "Also submitted ICCPR derogation",
      "States that derogated from the ECHR",
      "Also submitted ICCPR derogation"
    )
  ) %>% 
  tt(
    width = c(0.3, 0.2, 0.3, 0.2),
    notes = "ACHR = American Convention on Human Rights; ECHR = European Convention on Human Rights"
  ) %>% 
  format_tt(replace = list(" " = NA)) %>% 
  style_tt(j = 1:4, align = "lclc") %>% 
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = "cells={font=\\scriptsize}, row{1}={valign=b}"
  )
```

\clearpage

```{r tbl-results-determinants}
#| tbl-cap: Results from Authors (2024) showing the determinants of derogation
#| include: true
#| echo: false

det_models <- readRDS(here::here("data", "raw_data",  "determinants_models.rds"))

modelsummary(
  det_models$models, 
  exponentiate = TRUE,
  estimate = "{estimate} ({std.error}){stars}",
  statistic = NULL,
  coef_map = det_models$coef_map,
  gof_map = det_models$gof_map,
  output = "tinytable",
  fmt = fmt_significant(2),
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  notes = list(
    "Dependent variable = derogation submission",
    "Logistic regression models; odds ratios (standard error)",
    "* p < 0.1, ** p < 0.05, *** p < 0.01"
  ),
  width = c(0.5, 0.25, 0.25)
) %>% 
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = "cells={font=\\scriptsize}"
  )
```

\clearpage

# References