### 2024-11-26 README.qmd
Add official DOI

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)

manuscript_target <- targets::tar_read(manuscript)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
```

# `r title`

[Suparna Chaudhry](https://www.suparnachaudhry.com/) • Lewis and Clark College  
[Audrey Comstock](https://audreylcomstock.weebly.com/) • Arizona State University  
[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

---

<!-- badges: start -->
[![Global Policy DOI](https://img.shields.io/badge/International%20Interactions%20DOI-10.1080%2F03050629.2024.2413965-brightgreen)](https://doi.org/10.1080/03050629.2024.2413965) [![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fy8ked-blue)](https://doi.org/10.31235/osf.io/y8ked) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FAQVNK-blue)](https://doi.org/10.17605/OSF.IO/AQVNK) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
<!-- badges: end -->

> Suparna Chaudhry, Audrey L. Comstock, and Andrew Heiss, “Pandemic Pass? Treaty Derogations and Human Rights Practices during COVID-19,” *International Interactions* 50, no. 6 (2024): 1064–86, <https://doi.org/10.1080/03050629.2024.2413965>.

---

## Abstract

`r abstract`

---

## How to download and replicate

You can either [download the compendium as a ZIP file](/archive/main.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

We use the [{renv} package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [{targets} package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis. ([See this for a short helpful walkthrough of {targets}.](https://books.ropensci.org/targets/walkthrough.html)).

To replicate the findings and re-run the analysis, you can use a Docker container at [`mountainous-mackerel-docker`](https://github.com/andrewheiss/mountainous-mackerel-docker), which will create a complete computing envirionment for running everything. Full details and instructions [are available there](https://github.com/andrewheiss/mountainous-mackerel-docker).

Alternatively, everything can be run on your local computer too. Do the following:

1. Download and install these fonts (if you’re using Windows, make sure you right click on the font files and choose “Install for all users” when installing these fonts):
    - [Noto Sans](https://fonts.google.com/specimen/Noto+Sans)
    - [Linux Libertine O](https://www.cufonfonts.com/font/linux-libertine-o) (also [here](https://sourceforge.net/projects/linuxlibertine/))
    - [Libertinus Math](https://github.com/alerque/libertinus)

2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library. Also [install Homebrew](https://brew.sh/) and run `brew install ghostscript`.
    - If you’re using Windows, [install RTools too](https://cran.r-project.org/bin/windows/Rtools/) and add it to your PATH so that you can install packages from source if needed

3. Open `mountainous-mackerel.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).

4. In the terminal, run `quarto install tinytex` to ensure that you have a working LaTeX installation.

5. If it's not installed already, R *should* try to install the {renv} package when you open the RStudio Project for the first time. If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.

6. Run this to install {cmdstanr}. This is supposed to happen automatically as part of `renv::restore()` below, since {cmdstanr} is in the lockfile, but [due to an issue with {renv}](https://github.com/rstudio/renv/issues/1961) (fixed in the development version as of 2024-08-06), it doesn't install correctly because it is hosted at <https://stan-dev.r-universe.dev> instead of CRAN. So for now, until the next stable release of {renv}, it's easiest to install {cmdstanr} in a separate step.

   ```r
   install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", "https://packagemanager.posit.co/cran/latest"))
   ```

7. Run `renv::restore()` in the R console to install all the required packages for this project.

8. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.

Running `targets::tar_make()` will create several helpful outputs:

1. All project data in `data/`
2. An analysis notebook website in `_site/`
3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/`

## `r emoji_glue(":snow_capped_mountain::fish:")`: Note on "mountainous mackerel" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.4.0

codename(seed = "covid and ngos", type = "ubuntu")
#> [1] "mountainous mackerel"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.